### 2024-11-19 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@scams-research}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2025-01-20 Fig1.py
Added python script for figure

@@ -1,0 +1,309 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import MDAnalysis as mda
+from MDAnalysis.coordinates.memory import MemoryReader
+from MDAnalysis.transformations import wrap 
+from MDAnalysis.transformations.boxdimensions import set_dimensions
+import matplotlib.gridspec as gridspec
+import matplotlib 
+from matplotlib.patches import Arc
+
+
+# Function definitions 
+
+def pi_centrer(coords,weights,box_min, box_width):
+    """
+    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+    frac_coords = (coords - box_min) / box_width
+    theta = frac_coords * (2 * np.pi) 
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    xi_bar = np.average(xi,weights=weights)
+    zeta_bar = np.average(zeta,weights=weights)
+    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
+    new_s_coords = (theta_bar) / (2 * np.pi)
+    new_s_coords = new_s_coords*box_width + box_min
+
+    return new_s_coords
+
+
+def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
+    """
+    Calculates the xi and zeta used within the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+
+    frac_coords = (coords - box_min) / box_width 
+    theta = frac_coords * (2 * np.pi) + np.pi
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    return xi, zeta
+
+def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
+    """
[…]

### 2025-01-20 Fig2.py
Code for Figure 2

@@ -1,0 +1,204 @@
+
+import MDAnalysis as mda
+from MDAnalysis.coordinates.memory import MemoryReader
+from MDAnalysis.transformations import wrap 
+from MDAnalysis.transformations.boxdimensions import set_dimensions
+import numpy as np
+import matplotlib.pyplot as plt
+from MDAnalysis.transformations.boxdimensions import set_dimensions
+import matplotlib.gridspec as gridspec
+import matplotlib as matplotlib
+import matplotlib.patches as patches
+
+# Function definitions 
+def pi_centrer(coords,weights,box_min, box_width):
+    """
+    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+    frac_coords = (coords - box_min) / box_width
+    theta = frac_coords * (2 * np.pi) 
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    xi_bar = np.average(xi,weights=weights)
+    zeta_bar = np.average(zeta,weights=weights)
+    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
+    new_s_coords = (theta_bar) / (2 * np.pi)
+    new_s_coords = new_s_coords*box_width + box_min
+
+    return new_s_coords
+
+
+def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
+    """
+    Calculates the xi and zeta used within the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+
+    frac_coords = (coords - box_min) / box_width 
+    theta = frac_coords * (2 * np.pi) + np.pi
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    return xi, zeta, theta
+
+def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
[…]

### 2025-01-20 Fig3.py
Added Figure 3 python file

@@ -1,0 +1,102 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+from scipy.stats import uniform
+import matplotlib as matplotlib
+
+# Function definitions 
+def pi_centrer(coords,weights,box_min, box_width):
+    """
+    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+    frac_coords = (coords - box_min) / box_width
+    theta = frac_coords * (2 * np.pi) 
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    xi_bar = np.average(xi,weights=weights)
+    zeta_bar = np.average(zeta,weights=weights)
+    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
+    new_s_coords = (theta_bar) / (2 * np.pi)
+    new_s_coords = new_s_coords*box_width + box_min
+
+    return new_s_coords
+
+# Main
+m = int(2 ** 24) 
+k = np.zeros((m, 6))
+x_max = 1
+x_min = 0
+
+# Variables to define wrapping
+minimum_uniform = 0.65
+
+
+for i in tqdm(range(m)):
+    rng = np.random.default_rng(i)
+    N = rng.integers(3, 512, size=1)
+    n = rng.uniform(0.1, 0.45, size=1)
+    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
+    y = particles
+    particle_span = particles.max() - particles.min()
+    y_pbc = y % x_max
+    k[i, 0] = N
+    k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
+    k[i, 2] = np.average(y)
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y_pbc - offset) % (x_max - x_min)
+    k[i, 3] = np.average(moved) + offset
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y - offset) % (x_max - x_min)
+    # aymmetry metric
[…]

### 2025-01-21 src/scripts/Fig1.py
cleared up code and added imports

@@ -1,152 +1,19 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import MDAnalysis as mda
-from MDAnalysis.coordinates.memory import MemoryReader
-from MDAnalysis.transformations import wrap 
-from MDAnalysis.transformations.boxdimensions import set_dimensions
 import matplotlib.gridspec as gridspec
 import matplotlib 
 from matplotlib.patches import Arc
+from Functions import pi_centrer,test_pi_centrer,xi_zeta_calc,xi_zeta_bar_calc,draw_dashed_box
+from Variables import * 
 
 
-# Function definitions 
-
-def pi_centrer(coords,weights,box_min, box_width):
-    """
-    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
-    frac_coords = (coords - box_min) / box_width
-    theta = frac_coords * (2 * np.pi) 
-    xi = np.cos(theta)
-    zeta = np.sin(theta)
-    xi_bar = np.average(xi,weights=weights)
-    zeta_bar = np.average(zeta,weights=weights)
-    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
-    new_s_coords = (theta_bar) / (2 * np.pi)
-    new_s_coords = new_s_coords*box_width + box_min
-
-    return new_s_coords
-
-
-def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
-    """
-    Calculates the xi and zeta used within the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
-
-    frac_coords = (coords - box_min) / box_width 
-    theta = frac_coords * (2 * np.pi) + np.pi
-    xi = np.cos(theta)
-    zeta = np.sin(theta)
-    return xi, zeta
-
-def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
-    """
[…]

### 2025-01-21 src/scripts/Fig2.py

@@ -9,114 +9,8 @@ from MDAnalysis.transformations.boxdimensions import set_dimensions
 import matplotlib.gridspec as gridspec
 import matplotlib as matplotlib
 import matplotlib.patches as patches
-
-# Function definitions 
-def pi_centrer(coords,weights,box_min, box_width):
-    """
-    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
-    frac_coords = (coords - box_min) / box_width
-    theta = frac_coords * (2 * np.pi) 
-    xi = np.cos(theta)
-    zeta = np.sin(theta)
-    xi_bar = np.average(xi,weights=weights)
-    zeta_bar = np.average(zeta,weights=weights)
-    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
-    new_s_coords = (theta_bar) / (2 * np.pi)
-    new_s_coords = new_s_coords*box_width + box_min
-
-    return new_s_coords
-
-
-def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
-    """
-    Calculates the xi and zeta used within the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
-
-    frac_coords = (coords - box_min) / box_width 
-    theta = frac_coords * (2 * np.pi) + np.pi
-    xi = np.cos(theta)
-    zeta = np.sin(theta)
-    return xi, zeta, theta
-
-def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
-    """
-    Calculates the xi bar, zeta bar and theta bar used within the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
[…]

### 2025-01-21 src/scripts/Fig3.py

@@ -3,30 +3,7 @@ import matplotlib.pyplot as plt
 from tqdm import tqdm
 from scipy.stats import uniform
 import matplotlib as matplotlib
-
-# Function definitions 
-def pi_centrer(coords,weights,box_min, box_width):
-    """
-    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
-    
-    Parameters:
-        coords (np.array): 1d array of particle coordinates
-        weights (np.array): 1d array of particle masses
-        box_min (float): Minimum x-coordinate of the box.
-        box_width (float): Width of the box
-
-    """
-    frac_coords = (coords - box_min) / box_width
-    theta = frac_coords * (2 * np.pi) 
-    xi = np.cos(theta)
-    zeta = np.sin(theta)
-    xi_bar = np.average(xi,weights=weights)
-    zeta_bar = np.average(zeta,weights=weights)
-    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
-    new_s_coords = (theta_bar) / (2 * np.pi)
-    new_s_coords = new_s_coords*box_width + box_min
-
-    return new_s_coords
+from Functions import pi_centrer
 
 # Main
 m = int(2 ** 24) 

### 2025-01-21 src/scripts/Functions.py

@@ -1,0 +1,132 @@
+# Function definitions 
+import numpy as np
+import MDAnalysis as mda
+from MDAnalysis.coordinates.memory import MemoryReader
+from MDAnalysis.transformations import wrap 
+from MDAnalysis.transformations.boxdimensions import set_dimensions
+
+def pi_centrer(coords,weights,box_min, box_width):
+    """
+    Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+    frac_coords = (coords - box_min) / box_width
+    theta = frac_coords * (2 * np.pi) 
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    xi_bar = np.average(xi,weights=weights)
+    zeta_bar = np.average(zeta,weights=weights)
+    theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
+    new_s_coords = (theta_bar) / (2 * np.pi)
+    new_s_coords = new_s_coords*box_width + box_min
+
+    return new_s_coords
+
+
+def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
+    """
+    Calculates the xi and zeta used within the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
+
+    frac_coords = (coords - box_min) / box_width 
+    theta = frac_coords * (2 * np.pi) + np.pi
+    xi = np.cos(theta)
+    zeta = np.sin(theta)
+    return xi, zeta
+
+def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
+    """
+    Calculates the xi bar, zeta bar and theta bar used within the bai and breen method.
+    
+    Parameters:
+        coords (np.array): 1d array of particle coordinates
+        weights (np.array): 1d array of particle masses
+        box_min (float): Minimum x-coordinate of the box.
+        box_width (float): Width of the box
+
+    """
[…]

### 2025-01-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aip]{revtex4-2}
 \usepackage{showyourwork}
+\usepackage[version=4]{mhchem}
+\usepackage{graphicx}
+\usepackage[utf8]{inputenc}
+\graphicspath{{figures/}}
+\usepackage{bm}
+\usepackage{setspace}
+\usepackage{siunitx}
+\sisetup{list-units=bracket,
+         list-final-separator={, and },
+         range-units=bracket,
+         range-phrase={, },
+         uncertainty-mode=separate
+         }
+\usepackage{lipsum}
+\newcommand{\papertitle}{On the Estimation of Centre of Mass in Periodic Systems}
+\usepackage[noabbrev,nameinlink,capitalize]{cleveref}
+\crefname{equation}{Eq.}{Eqs.}
+\crefformat{equation}{Eq.~#2#1#3}
+\crefrangeformat{equation}{Eqs.~#3#1#4 to~#5#2#6}
+\crefmultiformat{equation}{Eqs.~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3}
+\crefname{figure}{Fig.}{Figs.}
 
-% Begin!
+%% Apr 2021: AIP requests that the corresponding 
+%% email to be moved after the affiliations
+\makeatletter
+\def\@email#1#2{%
+ \endgroup
+ \patchcmd{\titleblock@produce}
+  {\frontmatter@RRAPformat}
+  {\frontmatter@RRAPformat{\produce@RRAP{*#1\href{mailto:#2}{#2}}}\frontmatter@RRAPformat}
+  {}{}
+}%
+\makeatother
+
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+\let\oldaddcontentsline\addcontentsline
+\renewcommand{\addcontentsline}[3]{}
 
-% Author list
-\author{@scams-research}
+%%% Potential Journals To Keep in Mind
+%% These all accept Letter style papers
+% Journal of Chemical Information and Modelling
+% Physical Review E
+% Journal of Chemical Physics
 
-% Abstract with filler text
+\title[Accurate Centre of Mass Estimation]{\papertitle}
+
+\author{Harry Richardson}
+  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
+\author{Josh Dunn}
+  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
+\author{Andrew R. McCluskey}
+\email{andrew.mccluskey@bristol.ac.uk}
+  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
+  \affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}
+
+\date{\today}
+
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+Calculation of the centre of mass of some species from a periodic cell is an important aspect of chemical and physical simulation. 
+One popular approach, described by Bai and Breen calculates the centre of mass via the projection of the individual particles coordinates on to a circle.
+This approach suffers from some numerical error as it is mathematically equivalent to finding the first moment of the Fourier series of the mass density. 
+Here we discuss this inaccuracy and propose an extension that overcomes it, enabling improved accuracy across computational simulation. 
 \end{abstract}
 
-% Main body with filler text
+\maketitle
+
 \section{Introduction}
-\label{sec:intro}
+\label{sec:intro}The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
+However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 
+Any group of particles will have the same number of valid centres of mass---positions in the simulation cell where the weighted relative positions of the masses sum to zero---as the number of particles (\cref{fig:problems}a).
+However, in the physical sciences, we may be concerned with the centre of mass of a single bonded molecule or a defined group of particles. 
+For a single bonded molecule there is only one correct answer for centre of mass.
+Practically speaking, for the majority of simulations, it is most desirable to compute the centre of mass that minimises the periodic distance to the measured particles.
+%
+\begin{figure*}
+  \includegraphics[width=\textwidth]{figures/full_banner2.pdf}
+  \caption{
+  (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where there are two periodic cells visible and the dashed line indicates the particle grouping that leads to a given centre of mass (cross).
+  (b) The problem with the naive centre of mass (red cross) calculation in periodic systems, where the particle group spans a periodic boundary. 
+  (c) The projection of the particles in the bottom diagram of (b) onto a circle to find the centre of mass, which minimises the weighted distance to all particles.}
+  \label{fig:problems}
+\end{figure*}
+%
+In a non-periodic system, the identification of the ``correct'' centre of mass of a group of particles, the average position weighted by the masses, is trivial. 
+However, where the periodic boundary can intersect the particle group, this naive calculation may be ``incorrect'' (\cref{fig:problems}cb). 
+This problem can be resolved where bonding or other relational information is present, but this is not always the case for computational simulation trajectories.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+Bai and Breen proposed an elegant algorithm to efficiently compute the minimized periodic distance centre of mass\cite{bai_calculating_2008}, involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
+However, this approach is approximate and introduces numerical error, regardless of whether particles span the periodic boundary. 
+Here, we propose an extension to the Bai and Breen algorithm removes this numerical error, allowing the correct centre of mass to be found in nearly all practical cases.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\section{Bai and Breen Approach}
 
+To outline the approach of Bai and Breen and the source of the numerical error, we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}. 
+This system is analogous to a fractional coordinate system. 
+A particle group of $N$ particles, can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
+Each particle $i$ is then projected onto the two dimensions of a unit circle, $\bm{\xi}$ and $\bm{\zeta}$, where, 
+%
+\begin{equation}
+    \xi_i = \cos(2\pi x_i),
+\end{equation}
+%
+and 
+%
+\begin{equation}
+    \zeta_i = \sin(2\pi x_i).
+\end{equation}
+% 
+The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses;
+%
+\begin{equation}
+    \bar{\xi} = \frac{1}{N}\sum_{i=1}^{N}m_i\xi_i,
+\end{equation}
+% 
+and
+%
+\begin{equation}
+    \bar{\zeta} = \frac{1}{N}\sum_{i=1}^{N}m_i\zeta_i.
+\end{equation}
+% 
+The projection process is then reversed for the weighted average circular coordinates.
+Practically, this is achieved using the 2-argument arctangent, 
+%
+\begin{equation}
+    \bar{x} = \frac{\text{atan}2(-\bar{\zeta}, -\bar{\xi}) + \pi}{2\pi},
+\end{equation}
+%
+to give the centre of mass position.
+In the case of an orthogonal multi-dimensional system, this method can be repeated independently for each dimension.
+
+While not discussed in the original work, this method is equivalent to finding the first moment of the Fourier series of the mass density as a function of position (see Appendix \ref{app:math}). 
+This approach is used in signal processing to find the spectral centroid of a signal\cite{tzanetakis_musical_2002}, where it is known that it is affected by noise and asymmetry\cite{teague_robust_2018}.
+Describing an asymmetric function with a single sine wave leads to an incorrect estimate of the spectral centroid, or centre of mass.
+
+Considering the limiting case of a 3 particle system, in which a central particle is moved between the 2 edge particles.
+The error in the center of mass varies, reaching its maximum very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}). 
+At the point where the particles are equally distributed, the system is perfectly symmetric and there is no numerical error in the centre of mass. 
+Similarly, there is no error in one or two particle systems as these are inherently symmetrical. 
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/degenerate_graph.pdf}
+    \caption{Quantification of the limiting case (bottom), a 3 particle molecule spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the molecule.
+    This error is shown visually for a single example (top), where the Bai and Breen estimate (orange cross) is not close to the true centre of mass (green cross).}
+    \label{fig:error_quantification}
+\end{figure}
+%
+
+\section{\emph{Pseudo}-Centre of Mass Recentering}
+
+We suggest an extension that removes the error introduced by asymmetry, which involves using the Bai and Breen estimate as a \emph{pseudo}-centre of mass. 
+The group of particles is repositioned to the middle of the simulation cell using the \emph{pseudo}-centre of mass, such that there is no risk that the group spans a periodic boundary condition. 
+Then, the standard weighted mean may be used to find the centre of mass, before restoring this to the original space.
+
+The use of this recentering assumes that the inaccurate Bai and Breen method is capable of finding a centre of mass that is close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
+This assumption holds true for all systems where the particle group is less than half the size of the periodic cell, and is often still true unless the system is very asymmetric. 
+If a group of particles spans more than half the size of the periodic box, the only way to guarantee the centre of mass is found accurately is through the inclusion of bonding or other relational information.
+
+\section{Method Comparison}
+
+To demonstrate the effectiveness of our \emph{psuedo}-centre of mass recentering extension, \cref{fig:method_comparison} compares it to the Bai and Breen method.
+To achieve this, we compute the difference between the estimated centre of mass and the true value, as a ratio of the span of the particle group. 
+This error is compared with the asymmetry of the particle mass density in the periodic cell\cite{xioajun_on_1991}.
+In \cref{fig:method_comparison} as the asymmetry increases, so does the error in the Bai and Breen approach. 
+However, the \emph{psuedo}-centre of mass resampling is not affected in this way. 
+Therefore, for the majority of chemical simulations, this \emph{pseudo}-center of mass recentering extension should be used to ensure accuracy in centre of mass calculation. 
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/Asymmetry.pdf}
+    \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
+    $2^24$ random configurations of one-dimensional periodic boundary spanning molecules were generated, with between 3 and 512 atoms and total length of less than half the box. 
+    For each configuration, the two approaches were used and the asymmetry of the molecule\cite{xioajun_on_1991} was compared to the error in the estimate normalised by the span (length) of the molecule.
+    The results have been binned along the the $x$-axis using 100 bins, spaced evenly on a log scale.}
+    \label{fig:method_comparison}
+    \script{Fig3}
+\end{figure}
+%
+
+\section{Conclusions}
+
+We have presented the \emph{psuedo}-centre of mass recentering extension to the Bai and Breen approach for the determination of the centre of mass of a group of particles in a periodic cell. 
+Unlike the existing approach, this method has no error, even at very large values of asymmetry.
+To enable the use of this approach, we have implemented it in the open-source Python package \textsc{kinisi}\cite{mccluskey_kinisi_2024} and hope that others will consider its use in future.
+
+\begin{acknowledgements}
+We thank Richard Gowers for the insightful discussion that started us down this rabbit hole.
+\end{acknowledgements}
+
+\section*{Conflict of Interest Statement}
+
+The authors have no conflicts to disclose
+
+\section*{Author Contributions}
+CRediT author statement: 
+H.R.: Formal analysis, Investigation, Methodology, Software, Visualization, Writing – original draft.
+J.D.: Conceptualization, Investigation, Methodology, Supervision, Writing – review \& editing.
+A.R.M.: Formal analysis, Funding acquisition, Investigation, Methodology, Project administration, Supervision, Writing – review \& editing.
+
+\section*{Data Availability}
+
+The data that support the findings of this study are openly available in a GitHub repository\cite{richardson_github_2025}, under an MIT license, including a complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}\cite{luger_showyourwork_2021}.
+
+\appendix
+
+\section{Mathematical Equivalency of Bai and Breen and First Moment Methods}
+\label{app:math}
+
+This appendix presents the mathematical equivalency between the Bai and Breen method for calculating the centre of mass, and finding the maximum of the first term of the Fourier series of the mass density as a function of position (known as the first moment method\cite{teague_robust_2018}).
+
+For a real valued function, the Fourier series can be described as 
+%
+\begin{equation}
+    s_N(x) = a_0 + \sum_{n=1}^{N} \left[ a_n \cos \left( 2\pi \frac{n}{P} x \right) + b_n \sin \left( 2\pi \frac{n}{P} x \right) \right],
+    \label{equ:fseries}
+\end{equation}
+%
+where, $N$ is some finite number of terms that are computed and $P$ is the period of the function. 
+Eqn.~\ref{equ:fseries} is equivalent to
+%
+\begin{equation}
+    s_N(x) = A_0 + \sum_{n=1}^{N} A_n \cos\left(2\pi \frac{n}{P}x - \varphi_n\right),
+\end{equation}
+%
+by the identity, 
+%
+\begin{equation}
+    \begin{aligned}
+        \cos \left( 2\pi \frac{n}{P} x - \varphi_n \right) = & \cos(\varphi_n) \cos \left( 2\pi \frac{n}{P} x \right)  \\ 
+        & + \sin(\varphi_n) \sin \left( 2\pi \frac{n}{P} x \right),
+    \end{aligned}
+\end{equation}
+where $\varphi_n$ is the phase shift of the $n$-th harmonic. 
+If we write, 
+%
+\begin{equation}
+    a_n = A_n \cos(\varphi_n),
+\end{equation}
+%
+and
+%
+\begin{equation}
+    b_n = A_n \sin(\varphi_n).
+\end{equation}
+%
+Then $a_n$ and $b_n$ can be considered to be constants that ``weight'' the relative contributions of $\sin \left( 2\pi \dfrac{n}{P} x \right)$ and $\cos \left( 2\pi \dfrac{n}{P} x \right)$, akin to the weighted average of $\bm{\xi}$ and $\bm{\zeta}$ carried out within the B\&B method.
+
+As stated in the main text, the Bai and Breen method is equivalent to finding the maxima of the first function of the Fourier series, $f_1$,
+%
+\begin{equation}
+    f_1(x) = a_1 \cos{\Big(2\pi\frac{1}{P} x\Big)} + b_1 \sin{\Big(2\pi\frac{1}{P} x\Big)},
+\end{equation} 
+%
+where $P$ is $1$ as the Bai and Breen method projects the full box size on to a single unit circle. 
+To maximise the value of $f_1(x)$ a stationary point is found where,
+%
+\begin{equation}
+    \frac{\partial f_1(x)}{\partial x} = 0,
+\end{equation}
+%
+which is 
+%
+\begin{equation}
+    \frac{\partial f_1}{\partial x} = -2 \pi a_1 \sin( 2\pi x) + 2 \pi b_1 \cos( 2\pi x) = 0.
+    \label{equ:der}
+\end{equation}
+%
+Eqn.~\ref{equ:der} can be rearranged to give, 
+%
+\begin{equation}
+\frac{\sin(2 \pi x)}{\cos(2 \pi x)} =  \frac{b_1}{a_1}
+\end{equation}
+%
+and from the trigonometric identity, $\tan(x) = \sin(x) / \cos(x)$, 
+%
+\begin{equation}
+    \tan(2 \pi x) = \frac{b_1}{a_1}.
+\end{equation}
+%
+Therefore the position at which the first function of the Fourier series is maximized, $x$, can be calculated as, 
+%
+\begin{equation}
+    x = \frac{\text{atan}2(b_1, a_1)}{2\pi}.
+\end{equation}
+%
+In terms of periodic position, 
+%
+\begin{equation}
+    \frac{\text{atan}2(b_1, a_1)}{2\pi} \equiv \frac{\text{atan}2(-b_1, -a_1) + \pi}{2\pi}.
+\end{equation}
+%
+The negative $b_1$ and $a_1$ in the Bai and Breen method ensure that the centre of mass is always in the correct periodic box with no additional wrapping.
+
+\nocite{*}
 \bibliography{bib}

### 2025-01-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\section{Introduction}
-\label{sec:intro}The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
+% \section{Introduction}
+% \label{sec:intro}The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
 However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 

### 2025-01-21 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2025-01-21 src/tex/ms.tex
@@
 \section{Introduction}
-\label{sec:intro}The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
+\label{sec:intro}Th% e calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
 However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 

### 2025-01-21 src/scripts/Fig1.py
Savefigs

@@ -5,7 +5,7 @@ import matplotlib
 from matplotlib.patches import Arc
 from Functions import pi_centrer,test_pi_centrer,xi_zeta_calc,xi_zeta_bar_calc,draw_dashed_box
 from Variables import * 
-
+import paths
 
 
 fig = plt.figure(figsize=(8.06, 8.06/4.6666667))
@@ -165,4 +165,5 @@ ax2.axis('off')
 fig.text(.30, .05, '(a)', ha='center', size = font_size)
 fig.text(.605, .05, '(b)', ha='center', size = font_size)
 fig.text(.818, .05, '(c)', ha='center', size = font_size)
-plt.show()
+plt.savefig(paths.figures / 'Fig3.pdf')
+plt.close()

### 2025-01-21 src/scripts/Fig2.py

@@ -11,6 +11,7 @@ import matplotlib as matplotlib
 import matplotlib.patches as patches
 from Functions import pi_centrer, xi_zeta_calc, xi_zeta_bar_calc, test_pi_centrer
 from Variables import * 
+import paths
 
 # Main
 
@@ -89,4 +90,5 @@ ax0.scatter(-20,-20, s=s_set, color = '#029E73', marker = 'x', label = 'True COM
 
 
 ax0.spines['top'].set_visible(False)
-plt.show()
+plt.savefig(paths.figures / 'Fig3.pdf')
+plt.close()

### 2025-01-21 src/scripts/Fig3.py

@@ -4,6 +4,7 @@ from tqdm import tqdm
 from scipy.stats import uniform
 import matplotlib as matplotlib
 from Functions import pi_centrer
+import paths
 
 # Main
 m = int(2 ** 24) 
@@ -76,4 +77,4 @@ for i, ci in enumerate(credible_intervals):
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()
-plt.show()
+plt.savefig(paths.figures / 'Fig3.pdf')

### 2025-01-21 src/scripts/Fig1.py
Fix filename

@@ -165,5 +165,5 @@ ax2.axis('off')
 fig.text(.30, .05, '(a)', ha='center', size = font_size)
 fig.text(.605, .05, '(b)', ha='center', size = font_size)
 fig.text(.818, .05, '(c)', ha='center', size = font_size)
-plt.savefig(paths.figures / 'Fig3.pdf')
+plt.savefig(paths.figures / 'Fig1.pdf')
 plt.close()

### 2025-01-21 src/scripts/Fig2.py
Fix filename

@@ -90,5 +90,5 @@ ax0.scatter(-20,-20, s=s_set, color = '#029E73', marker = 'x', label = 'True COM
 
 
 ax0.spines['top'].set_visible(False)
-plt.savefig(paths.figures / 'Fig3.pdf')
+plt.savefig(paths.figures / 'Fig2.pdf')
 plt.close()

### 2025-01-21 src/code/sampling.py
Spliting code and plotting

@@ -1,0 +1,44 @@
+import numpy as np
+from tqdm import tqdm
+from scipy.stats import uniform
+from Functions import pi_centrer
+import paths
+
+# Main
+m = int(2 ** 24) 
+k = np.zeros((m, 6))
+x_max = 1
+x_min = 0
+
+# Variables to define wrapping
+minimum_uniform = 0.65
+
+
+for i in tqdm(range(m)):
+    rng = np.random.default_rng(i)
+    N = rng.integers(3, 512, size=1)
+    n = rng.uniform(0.1, 0.45, size=1)
+    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
+    y = particles
+    particle_span = particles.max() - particles.min()
+    y_pbc = y % x_max
+    k[i, 0] = N
+    k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
+    k[i, 2] = np.average(y)
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y_pbc - offset) % (x_max - x_min)
+    k[i, 3] = np.average(moved) + offset
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y - offset) % (x_max - x_min)
+    # aymmetry metric
+    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
+    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
+    k[i, 4] = np.sum(np.abs(asymmetry))
+    k[i,5] = particle_span
+
+# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
+k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
+k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
+k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
+
+np.savetxt(paths.data / 'Fig3.npy', k)

### 2025-01-21 src/scripts/Fig3.py

@@ -1,48 +1,9 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from tqdm import tqdm
-from scipy.stats import uniform
 import matplotlib as matplotlib
-from Functions import pi_centrer
 import paths
 
-# Main
-m = int(2 ** 24) 
-k = np.zeros((m, 6))
-x_max = 1
-x_min = 0
-
-# Variables to define wrapping
-minimum_uniform = 0.65
-
-
-for i in tqdm(range(m)):
-    rng = np.random.default_rng(i)
-    N = rng.integers(3, 512, size=1)
-    n = rng.uniform(0.1, 0.45, size=1)
-    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
-    y = particles
-    particle_span = particles.max() - particles.min()
-    y_pbc = y % x_max
-    k[i, 0] = N
-    k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
-    k[i, 2] = np.average(y)
-    offset = (k[i, 1] - ((x_max - x_min) / 2))
-    moved = (y_pbc - offset) % (x_max - x_min)
-    k[i, 3] = np.average(moved) + offset
-    offset = (k[i, 1] - ((x_max - x_min) / 2))
-    moved = (y - offset) % (x_max - x_min)
-    # aymmetry metric
-    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
-    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
-    k[i, 4] = np.sum(np.abs(asymmetry))
-    k[i,5] = particle_span
-
-# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
-k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
-k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
-
+k = np.load(paths.data / 'Fig3.npy')
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100
@@ -78,3 +39,4 @@ axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()
 plt.savefig(paths.figures / 'Fig3.pdf')
+plt.close()

### 2025-01-21 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2025-01-21 src/tex/ms.tex
@@
 
-%%% Potential Journals To Keep in Mind
-%% These all accept Letter style papers
-% Journal of Chemical Information and Modelling
-% Physical Review E
-% Journal of Chemical Physics
-
 \title[Accurate Centre of Mass Estimation]{\papertitle}
@@
 \section{Introduction}
-\label{sec:intro}The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics~\\citehang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
+\label{sec:intro}
+
+% The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics\cite{zhang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
 However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 
@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{figures/full_banner2.pdf}
+  \includegraphics[width=\textwidth]{Fig1.pdf}
   \caption{
@@
   \label{fig:problems}
+  \script{Fig1.py}
 \end{figure*}
 %
+
 In a non-periodic system, the identification of the ``correct'' centre of mass of a group of particles, the average position weighted by the masses, is trivial. 
-However, where the periodic boundary can intersect the particle group, this naive calculation may be ``incorrect'' (\cref{fig:problems}cb). 
+However, where the periodic boundary can intersect the particle group, this naive calculation may be ``incorrect'' (\cref{fig:problems}b). 
 This problem can be resolved where bonding or other relational information is present, but this is not always the case for computational simulation trajectories.
@@
     \centering
-    \includegraphics[width=8cm]{figures/degenerate_graph.pdf}
+    \includegraphics[width=8cm]{Fig2.pdf}
     \caption{Quantification of the limiting case (bottom), a 3 particle molecule spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the molecule.
@@
     \label{fig:error_quantification}
+    \script{Fig2.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=8cm]{figures/Asymmetry.pdf}
+    \includegraphics[width=8cm]{Fig3.pdf}
     \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
@@
     \label{fig:method_comparison}
-    \script{Fig3}
+    \script{Fig3.py}
 \end{figure}

### 2025-01-21 src/scripts/Fig3.py
Allow pickle

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib as matplotlib
 import paths
 
-k = np.load(paths.data / 'Fig3.npy')
+k = np.load(paths.data / 'Fig3.npy', allow_pickle=True)
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-01-21 src/scripts/Fig3.py
Allow pickle

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib as matplotlib
 import paths
 
-k = np.load(paths.data / 'Fig3.npy', allow_pickle=True)
+k = np.loadtxt(paths.data / 'Fig3.npy')
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-01-22 src/scripts/Fig1.py
Playing with padding

@@ -165,5 +165,5 @@ ax2.axis('off')
 fig.text(.30, .05, '(a)', ha='center', size = font_size)
 fig.text(.605, .05, '(b)', ha='center', size = font_size)
 fig.text(.818, .05, '(c)', ha='center', size = font_size)
-plt.savefig(paths.figures / 'Fig1.pdf')
+plt.savefig(paths.figures / 'Fig1.pdf', bbox_inches='tight', pad_inches = 0)
 plt.close()

### 2025-01-22 src/scripts/Fig2.py

@@ -90,5 +90,5 @@ ax0.scatter(-20,-20, s=s_set, color = '#029E73', marker = 'x', label = 'True COM
 
 
 ax0.spines['top'].set_visible(False)
-plt.savefig(paths.figures / 'Fig2.pdf')
+plt.savefig(paths.figures / 'Fig2.pdf', bbox_inches='tight', pad_inches = 0))
 plt.close()

### 2025-01-22 src/scripts/Fig3.py

@@ -38,5 +38,5 @@ for i, ci in enumerate(credible_intervals):
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()
-plt.savefig(paths.figures / 'Fig3.pdf')
+plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches = 0))
 plt.close()

### 2025-01-22 src/scripts/Fig2.py
Playing with padding

@@ -90,5 +90,5 @@ ax0.scatter(-20,-20, s=s_set, color = '#029E73', marker = 'x', label = 'True COM
 
 
 ax0.spines['top'].set_visible(False)
-plt.savefig(paths.figures / 'Fig2.pdf', bbox_inches='tight', pad_inches = 0))
+plt.savefig(paths.figures / 'Fig2.pdf', bbox_inches='tight', pad_inches = 0)
 plt.close()

### 2025-01-22 src/scripts/Fig3.py

@@ -38,5 +38,5 @@ for i, ci in enumerate(credible_intervals):
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()
-plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches = 0))
+plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches = 0)
 plt.close()

### 2025-01-22 src/scripts/Fig2.py
Ticks

@@ -44,7 +44,6 @@ xmin = 4
 xmax = 16
 s_set = 50
 
-
 # Tick removal, horizontal line plotting
 for x in [ax1]:
     x.set_xlim(4,16.1)
@@ -56,7 +55,6 @@ for x in [ax1]:
     x.set_xticks([])
     x.axis('off')
 
-
 # Selecting for specific molecule
 particles = 3
 coords = np.array([xrange[0],xrange[-1],xrange[20]])
@@ -66,11 +64,10 @@ ax1.scatter(coords,np.zeros(particles), s = s_set, color = ['black','black','#01
 ax1.plot(coords,np.zeros(particles), color = 'grey', alpha = 0.8, linewidth = l_set)
 ax1.scatter(MDA_store[20],0, s=s_set, color = '#029E73', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
 ax1.scatter(pi_store[20],0, s=s_set, color = '#D55E00', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
-
+ax1.set_yticks([0, 0.1, 0.2, 0.3])
+ax1.set_xticks([0.25, 0.5, 0.75])
 
 ax0 = fig.add_subplot(gs[1,0])
-
-
 
 ax0.plot(xrange/20,np.abs(pi_store - MDA_store)/10, label = 'moving particle', color = '#0173B2',solid_capstyle='round')
 ax0.set_xlabel('Particle coordinates')
@@ -78,16 +75,12 @@ ax0.set_ylabel('Centre of mass error / Span of molecule')
 ax0.set_ylim(0,0.35)
 ax0.set_xlim(0.2,0.8)
 
-
-
 # Demonstration line
 ax1.axvline(xrange[20], ymax = 0.5, color = 'orange', linestyle = 'dashed')
 ax0.axvline(xrange[20]/20-0.0013, ymin = (np.abs(pi_store - MDA_store)/10)[20]/0.35, color = 'orange', linestyle = 'dashed')
 ax0.axhline((np.abs(pi_store - MDA_store)/10)[20], xmax = 0.25, color = 'orange', linestyle = 'dashed', label = 'example error')
[…]

### 2025-01-22 src/scripts/Fig2.py
Ticks

@@ -64,8 +64,6 @@ ax1.scatter(coords,np.zeros(particles), s = s_set, color = ['black','black','#01
 ax1.plot(coords,np.zeros(particles), color = 'grey', alpha = 0.8, linewidth = l_set)
 ax1.scatter(MDA_store[20],0, s=s_set, color = '#029E73', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
 ax1.scatter(pi_store[20],0, s=s_set, color = '#D55E00', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
-ax1.set_yticks([0, 0.1, 0.2, 0.3])
-ax1.set_xticks([0.25, 0.5, 0.75])
 
 ax0 = fig.add_subplot(gs[1,0])
 
@@ -74,6 +72,8 @@ ax0.set_xlabel('Particle coordinates')
 ax0.set_ylabel('Centre of mass error / Span of molecule')
 ax0.set_ylim(0,0.35)
 ax0.set_xlim(0.2,0.8)
+ax0.set_yticks([0, 0.1, 0.2, 0.3])
+ax0.set_xticks([0.25, 0.5, 0.75])
 
 # Demonstration line
 ax1.axvline(xrange[20], ymax = 0.5, color = 'orange', linestyle = 'dashed')

### 2025-01-22 src/scripts/Fig2.py
Plotting

@@ -45,15 +45,14 @@ xmax = 16
 s_set = 50
 
 # Tick removal, horizontal line plotting
-for x in [ax1]:
-    x.set_xlim(4,16.1)
-    x.set_ylim(-1,1)
-    x.hlines(y1, xmin, xmax, color = '#0173B2', alpha = 0.4, linewidth = blue_set)
-    x.set_yticklabels([])
-    x.set_yticks([])
-    x.set_xticklabels([])
-    x.set_xticks([])
-    x.axis('off')
+ax1.set_xlim(4,16.1)
+ax1.set_ylim(-1,1)
+ax1.hlines(y1, xmin, xmax, color = '#0173B2', alpha = 0.4, linewidth = blue_set)
+ax1.set_yticklabels([])
+ax1.set_yticks([])
+ax1.set_xticklabels([])
+ax1.set_xticks([])
+ax1.axis('off')
 
 # Selecting for specific molecule
 particles = 3
@@ -67,7 +66,7 @@ ax1.scatter(pi_store[20],0, s=s_set, color = '#D55E00', marker = 'x', label = 'N
 
 ax0 = fig.add_subplot(gs[1,0])
 
-ax0.plot(xrange/20,np.abs(pi_store - MDA_store)/10, label = 'moving particle', color = '#0173B2',solid_capstyle='round')
+ax0.plot(xrange / 20, np.abs(pi_store - MDA_store) / 10, label = 'moving particle', color = '#0173B2', solid_capstyle='round')
 ax0.set_xlabel('Particle coordinates')
 ax0.set_ylabel('Centre of mass error / Span of molecule')
 ax0.set_ylim(0,0.35)
@@ -77,7 +76,7 @@ ax0.set_xticks([0.25, 0.5, 0.75])
 
 # Demonstration line
 ax1.axvline(xrange[20], ymax = 0.5, color = 'orange', linestyle = 'dashed')
-ax0.axvline(xrange[20]/20-0.0013, ymin = (np.abs(pi_store - MDA_store)/10)[20]/0.35, color = 'orange', linestyle = 'dashed')
+ax0.axvline(xrange[20] / 20-0.0013, ymin = (np.abs(pi_store - MDA_store)/10)[20]/0.35, color = 'orange', linestyle = 'dashed')
 ax0.axhline((np.abs(pi_store - MDA_store)/10)[20], xmax = 0.25, color = 'orange', linestyle = 'dashed', label = 'example error')
[…]

### 2025-01-22 src/scripts/Fig3.py

@@ -21,7 +21,7 @@ for z in range(0,N_bins):
 
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]
-matplotlib.rcParams.update({'font.size': 7})
+# matplotlib.rcParams.update({'font.size': 7})
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
 
 axis.plot(k[:, 4], (k[:, 3] - k[:, 2])/k[:,5], label='This Work', color = 'orange',solid_capstyle='round')

### 2025-01-22 src/scripts/Fig3.py
Plotting

@@ -35,6 +35,8 @@ for i, ci in enumerate(credible_intervals):
                     lw=0)
     
 
+axis.set_yticks([-0.1, 0, 0.1])
+axis.set_xticks([0, 50, 100])
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()

### 2025-01-22 src/scripts/Fig3.py
Plotting

@@ -21,7 +21,6 @@ for z in range(0,N_bins):
 
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]
-# matplotlib.rcParams.update({'font.size': 7})
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
 
 axis.plot(k[:, 4], (k[:, 3] - k[:, 2])/k[:,5], label='This Work', color = 'orange',solid_capstyle='round')
@@ -35,7 +34,7 @@ for i, ci in enumerate(credible_intervals):
                     lw=0)
     
 
-axis.set_yticks([-0.1, 0, 0.1])
+axis.set_yticks([-0.08, 0.04, 0, 0.04, 0.08])
 axis.set_xticks([0, 50, 100])
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")

### 2025-01-22 src/scripts/Fig3.py
Plotting

@@ -34,7 +34,7 @@ for i, ci in enumerate(credible_intervals):
                     lw=0)
     
 
-axis.set_yticks([-0.08, 0.04, 0, 0.04, 0.08])
+axis.set_yticks([-0.08, -0.04, 0, 0.04, 0.08])
 axis.set_xticks([0, 50, 100])
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
-    $2^24$ random configurations of one-dimensional periodic boundary spanning molecules were generated, with between 3 and 512 atoms and total length of less than half the box. 
+    $2^{24}$ random configurations of one-dimensional periodic boundary spanning molecules were generated, with between 3 and 512 atoms and total length of less than half the box. 
     For each configuration, the two approaches were used and the asymmetry of the molecule\cite{xioajun_on_1991} was compared to the error in the estimate normalised by the span (length) of the molecule.

### 2025-01-22 src/scripts/Fig1.py
Yapf

@@ -1,43 +1,54 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
-import matplotlib 
+import matplotlib
 from matplotlib.patches import Arc
-from Functions import pi_centrer,test_pi_centrer,xi_zeta_calc,xi_zeta_bar_calc,draw_dashed_box
-from Variables import * 
+from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_calc, draw_dashed_box
+from Variables import *
 import paths
 
-
-fig = plt.figure(figsize=(8.06, 8.06/4.6666667))
+fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
 
 # box coordinates and line
 xmin = 0
 xmax = 20
 
+gs = gridspec.GridSpec(6,
+                       5,
+                       width_ratios=[1.9, 0.18, 1.05, 0.18, 0.9],
+                       height_ratios=[1, 1, 1, 1, 1, 1],
+                       wspace=0,
+                       hspace=0.2)
 
-
-gs = gridspec.GridSpec(6, 5, width_ratios=[1.9,0.18,1.05,0.18, 0.9], height_ratios=[1,1,1,1,1,1], wspace=0, hspace=0.2)
-
-# Axes for fig 1b 
+# Axes for fig 1b
 ax0 = fig.add_subplot(gs[0:3, 2])  # Top-left
 ax1 = fig.add_subplot(gs[3:6, 2])  # Bottom-left
 
 # Axes for fig 1c
-ax2 = fig.add_subplot(gs[:, 4]) 
+ax2 = fig.add_subplot(gs[:, 4])
 
-# Axes for fig 1a 
-ax3 = fig.add_subplot(gs[0:2,0]) 
-ax4 = fig.add_subplot(gs[2:4,0])
-ax5 = fig.add_subplot(gs[4:6,0])
-
+# Axes for fig 1a
+ax3 = fig.add_subplot(gs[0:2, 0])
+ax4 = fig.add_subplot(gs[2:4, 0])
+ax5 = fig.add_subplot(gs[4:6, 0])
 
 # Remove tick labels and plot horizontal and plot blue lines
-for x in [ax0,ax1]:
-    x.set_xlim(-1,21)
-    x.set_ylim(-1,1)
[…]

### 2025-01-22 src/scripts/Fig2.py

@@ -1,7 +1,6 @@
-
 import MDAnalysis as mda
 from MDAnalysis.coordinates.memory import MemoryReader
-from MDAnalysis.transformations import wrap 
+from MDAnalysis.transformations import wrap
 from MDAnalysis.transformations.boxdimensions import set_dimensions
 import numpy as np
 import matplotlib.pyplot as plt
@@ -10,7 +9,7 @@ import matplotlib.gridspec as gridspec
 import matplotlib as matplotlib
 import matplotlib.patches as patches
 from Functions import pi_centrer, xi_zeta_calc, xi_zeta_bar_calc, test_pi_centrer
-from Variables import * 
+from Variables import *
 import paths
 
 # Main
@@ -19,25 +18,30 @@ MDA_store = np.zeros(100)
 pi_store = np.zeros(100)
 corrected_store = np.zeros(100)
 
-# range x moves through 
-xrange = np.linspace(5.0001,14.999,100)
-xrange[50] = 5.0001 + (14.999-5.001) / 2 # force part of the xrange to be the zero errror point
-masses = np.array((1,1,1))
+# range x moves through
+xrange = np.linspace(5.0001, 14.999, 100)
+xrange[50] = 5.0001 + (
+    14.999 - 5.001) / 2  # force part of the xrange to be the zero errror point
+masses = np.array((1, 1, 1))
 
 for i, x in enumerate(xrange):
 
     # Positions of particles in system
-    positions = np.array((5.0001,x,14.999))
+    positions = np.array((5.0001, x, 14.999))
 
-    loop_md,loop_pi,loop_cor = test_pi_centrer(masses, positions)
+    loop_md, loop_pi, loop_cor = test_pi_centrer(masses, positions)
     MDA_store[i] = loop_md
     pi_store[i] = loop_pi
     corrected_store[i] = loop_cor
 
-
-gs = gridspec.GridSpec(2, 1, width_ratios=[1], height_ratios=[1,5], wspace=0, hspace=0)
+gs = gridspec.GridSpec(2,
+                       1,
+                       width_ratios=[1],
+                       height_ratios=[1, 5],
+                       wspace=0,
+                       hspace=0)
 fig = plt.figure(figsize=(4.03, 3.3))
-ax1 = fig.add_subplot(gs[0,0])
+ax1 = fig.add_subplot(gs[0, 0])
 
[…]

### 2025-01-22 src/scripts/Fig3.py

@@ -5,39 +5,41 @@ import paths
 
 k = np.loadtxt(paths.data / 'Fig3.npy')
 
-# Binning Bai and Breen error for visulation 
+# Binning Bai and Breen error for visulation
 N_bins = 100
 bins = np.logspace(-1, 2, N_bins)
 binned_indices = np.digitize(k[:, 4], bins)
 
-error_bb_array = np.empty([len(bins),2000000])
+error_bb_array = np.empty([len(bins), 2000000])
 error_bb_array[:] = np.nan
-err = (k[:, 1] - k[:, 2])/k[:,5]
-for z in range(0,N_bins):
+err = (k[:, 1] - k[:, 2]) / k[:, 5]
+for z in range(0, N_bins):
     length = len(np.where(binned_indices == z)[0])
     if length != 0:
-        error_bb_array[z,0:length] = err[np.where(binned_indices == z)[0]]
-
+        error_bb_array[z, 0:length] = err[np.where(binned_indices == z)[0]]
 
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]
-fig, axis = plt.subplots(figsize = (4.03, 3.3))
+fig, axis = plt.subplots(figsize=(4.03, 3.3))
 
-axis.plot(k[:, 4], (k[:, 3] - k[:, 2])/k[:,5], label='This Work', color = 'orange',solid_capstyle='round')
+axis.plot(k[:, 4], (k[:, 3] - k[:, 2]) / k[:, 5],
+          label='This Work',
+          color='orange',
+          solid_capstyle='round')
 
 for i, ci in enumerate(credible_intervals):
     axis.fill_between(bins,
-                    *np.nanpercentile(error_bb_array,ci, axis=1),
-                    alpha=alpha[i],
-                    color='#0173B2',
-                    label= f'${i+1}\sigma$ - B&B method',
-                    lw=0)
-    
+                      *np.nanpercentile(error_bb_array, ci, axis=1),
+                      alpha=alpha[i],
+                      color='#0173B2',
+                      label=f'${i+1}\sigma$ - B&B method',
+                      lw=0)
 
 axis.set_yticks([-0.08, -0.04, 0, 0.04, 0.08])
 axis.set_xticks([0, 50, 100])
+axis.set_ylim(-0.082, 0.082)
 axis.set_xlabel('Asymmetry')
[…]

### 2025-01-22 src/scripts/Functions.py

@@ -1,11 +1,12 @@
-# Function definitions 
+# Function definitions
 import numpy as np
 import MDAnalysis as mda
 from MDAnalysis.coordinates.memory import MemoryReader
-from MDAnalysis.transformations import wrap 
+from MDAnalysis.transformations import wrap
 from MDAnalysis.transformations.boxdimensions import set_dimensions
 
-def pi_centrer(coords,weights,box_min, box_width):
+
+def pi_centrer(coords, weights, box_min, box_width):
     """
     Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
     
@@ -17,19 +18,19 @@ def pi_centrer(coords,weights,box_min, box_width):
 
     """
     frac_coords = (coords - box_min) / box_width
-    theta = frac_coords * (2 * np.pi) 
+    theta = frac_coords * (2 * np.pi)
     xi = np.cos(theta)
     zeta = np.sin(theta)
-    xi_bar = np.average(xi,weights=weights)
-    zeta_bar = np.average(zeta,weights=weights)
+    xi_bar = np.average(xi, weights=weights)
+    zeta_bar = np.average(zeta, weights=weights)
     theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
     new_s_coords = (theta_bar) / (2 * np.pi)
-    new_s_coords = new_s_coords*box_width + box_min
+    new_s_coords = new_s_coords * box_width + box_min
 
     return new_s_coords
 
 
-def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
+def xi_zeta_calc(masses, coords, box_min=0, box_width=20):
     """
     Calculates the xi and zeta used within the bai and breen method.
     
@@ -41,13 +42,14 @@ def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
 
     """
 
-    frac_coords = (coords - box_min) / box_width 
+    frac_coords = (coords - box_min) / box_width
     theta = frac_coords * (2 * np.pi) + np.pi
     xi = np.cos(theta)
     zeta = np.sin(theta)
     return xi, zeta
 
-def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
+
+def xi_zeta_bar_calc(masses, coords, box_min=0, box_width=20):
     """
     Calculates the xi bar, zeta bar and theta bar used within the bai and breen method.
     
[…]

### 2025-01-22 src/scripts/sampling.py

@@ -5,7 +5,7 @@ from Functions import pi_centrer
 import paths
 
 # Main
-m = int(2 ** 24) 
+m = int(2**24)
 k = np.zeros((m, 6))
 x_max = 1
 x_min = 0
@@ -13,12 +13,11 @@ x_min = 0
 # Variables to define wrapping
 minimum_uniform = 0.65
 
-
 for i in tqdm(range(m)):
     rng = np.random.default_rng(i)
     N = rng.integers(3, 512, size=1)
     n = rng.uniform(0.1, 0.45, size=1)
-    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
+    particles = rng.uniform(minimum_uniform, minimum_uniform + n, size=N)
     y = particles
     particle_span = particles.max() - particles.min()
     y_pbc = y % x_max
@@ -31,10 +30,13 @@ for i in tqdm(range(m)):
     offset = (k[i, 1] - ((x_max - x_min) / 2))
     moved = (y - offset) % (x_max - x_min)
     # aymmetry metric
-    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
-    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
+    y_hist, x_hist = np.histogram(moved,
+                                  bins=np.linspace(x_min, x_max, 101),
+                                  density=True)
+    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size /
+                                                           2):][::-1]
     k[i, 4] = np.sum(np.abs(asymmetry))
-    k[i,5] = particle_span
+    k[i, 5] = particle_span
 
 # Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
 k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max

### 2025-01-22 src/scripts/Fig2.py
Normalisation

@@ -96,7 +96,7 @@ ax0.plot(xrange / 20,
          color='#0173B2',
          solid_capstyle='round')
 ax0.set_xlabel('Particle coordinates')
-ax0.set_ylabel('Centre of mass error / Span of molecule')
+ax0.set_ylabel('Normalised centre of mass error')
 ax0.set_ylim(0, 0.35)
 ax0.set_xlim(0.2, 0.8)
 ax0.set_yticks([0, 0.1, 0.2, 0.3])

### 2025-01-22 src/scripts/Fig3.py

@@ -39,7 +39,7 @@ axis.set_yticks([-0.08, -0.04, 0, 0.04, 0.08])
 axis.set_xticks([0, 50, 100])
 axis.set_ylim(-0.082, 0.082)
 axis.set_xlabel('Asymmetry')
-axis.set_ylabel("Centre of mass error / Span of particles")
+axis.set_ylabel('Normalised centre of mass error')
 plt.tight_layout()
 plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches=0)
 plt.close()

### 2025-01-22 src/scripts/Variables.py
Adding appropriate dependencie

@@ -2,7 +2,7 @@
 blue_set = 0.5
 y1 = 0
 height = 1
-font_size = 7
+font_size = 8
 cross_size = 1.5
 marker_set = 15
 l_set = 2

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-% The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics\cite{zhang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}.
+The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics.\cite{zhang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}
 However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 
@@
 
-Bai and Breen proposed an elegant algorithm to efficiently compute the minimized periodic distance centre of mass\cite{bai_calculating_2008}, involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
+Bai and Breen proposed an elegant algorithm to efficiently compute the minimized periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
 However, this approach is approximate and introduces numerical error, regardless of whether particles span the periodic boundary. 
@@
 While not discussed in the original work, this method is equivalent to finding the first moment of the Fourier series of the mass density as a function of position (see Appendix \ref{app:math}). 
-This approach is used in signal processing to find the spectral centroid of a signal\cite{tzanetakis_musical_2002}, where it is known that it is affected by noise and asymmetry\cite{teague_robust_2018}.
+This approach is used in signal processing to find the spectral centroid of a signal,\cite{tzanetakis_musical_2002} where it is known that it is affected by noise and asymmetry.\cite{teague_robust_2018}
 Describing an asymmetric function with a single sine wave leads to an incorrect estimate of the spectral centroid, or centre of mass.
@@
 To achieve this, we compute the difference between the estimated centre of mass and the true value, as a ratio of the span of the particle group. 
-This error is compared with the asymmetry of the particle mass density in the periodic cell\cite{xioajun_on_1991}.
+This error is compared with the asymmetry of the particle mass density in the periodic cell.\cite{xioajun_on_1991}
 In \cref{fig:method_comparison} as the asymmetry increases, so does the error in the Bai and Breen approach. 
@@
 
-This appendix presents the mathematical equivalency between the Bai and Breen method for calculating the centre of mass, and finding the maximum of the first term of the Fourier series of the mass density as a function of position (known as the first moment method\cite{teague_robust_2018}).
+This appendix presents the mathematical equivalency between the Bai and Breen method for calculating the centre of mass, and finding the maximum of the first term of the Fourier series of the mass density as a function of position, known as the first moment method.\cite{teague_robust_2018}
 

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \makeatother
+
+\renewcommand{\refname}{References}
 
@@
     \includegraphics[width=8cm]{Fig2.pdf}
-    \caption{Quantification of the limiting case (bottom), a 3 particle molecule spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the molecule.
+    \caption{Quantification of the limiting case (bottom), a 3 particle group spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the particle group.
     This error is shown visually for a single example (top), where the Bai and Breen estimate (orange cross) is not close to the true centre of mass (green cross).}
@@
     \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
-    $2^{24}$ random configurations of one-dimensional periodic boundary spanning molecules were generated, with between 3 and 512 atoms and total length of less than half the box. 
-    For each configuration, the two approaches were used and the asymmetry of the molecule\cite{xioajun_on_1991} was compared to the error in the estimate normalised by the span (length) of the molecule.
+    $2^{24}$ random configurations of one-dimensional periodic boundary spanning particle groups were generated, with between 3 and 512 particles and total length of less than half the box. 
+    For each configuration, the two approaches were used and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
     The results have been binned along the the $x$-axis using 100 bins, spaced evenly on a log scale.}
@@
 
-The authors have no conflicts to disclose
+The authors have no conflicts to disclose.
 
@@
 
-\nocite{*}
 \bibliography{bib}

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Calculation of the centre of mass of some species from a periodic cell is an important aspect of chemical and physical simulation. 
-One popular approach, described by Bai and Breen calculates the centre of mass via the projection of the individual particles coordinates on to a circle.
+One popular approach, described by Bai and Breen, calculates the centre of mass via the projection of the individual particles' coordinates onto a circle.
 This approach suffers from some numerical error as it is mathematically equivalent to finding the first moment of the Fourier series of the mass density. 
@@
 The calculation of the centre of mass of a group of particles, such as in a molecule, is important across computational chemistry and physics.\cite{zhang_chemically_2024,happel_coordinated_2024,maggi_universality_2021,grillo_molecular_2023,bullerjahn_unwrapping_2023,jaeger-honz_systematic_2024}
-However, for systems with periodic boundary conditions, computing the centre of mass is non-trivial. 
+However, computing the centre of mass is non-trivial for systems with periodic boundary conditions. 
 Any group of particles will have the same number of valid centres of mass---positions in the simulation cell where the weighted relative positions of the masses sum to zero---as the number of particles (\cref{fig:problems}a).
 However, in the physical sciences, we may be concerned with the centre of mass of a single bonded molecule or a defined group of particles. 
-For a single bonded molecule there is only one correct answer for centre of mass.
-Practically speaking, for the majority of simulations, it is most desirable to compute the centre of mass that minimises the periodic distance to the measured particles.
+For a single bonded molecule, there is only one correct answer for the centre of mass.
+Practically speaking, for most simulations, it is most desirable to compute the centre of mass that minimises the periodic distance to the measured particles.
 %
@@
   \caption{
-  (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where there are two periodic cells visible and the dashed line indicates the particle grouping that leads to a given centre of mass (cross).
+  (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where two periodic cells are visible, and the dashed line indicates the particle grouping that leads to a given centre of mass (cross).
   (b) The problem with the naive centre of mass (red cross) calculation in periodic systems, where the particle group spans a periodic boundary. 
@@
 
-In a non-periodic system, the identification of the ``correct'' centre of mass of a group of particles, the average position weighted by the masses, is trivial. 
+In a non-periodic system, identifying the ``correct'' centre of mass of a group of particles, the average position weighted by the masses, is trivial. 
 However, where the periodic boundary can intersect the particle group, this naive calculation may be ``incorrect'' (\cref{fig:problems}b). 
@@
 
-Bai and Breen proposed an elegant algorithm to efficiently compute the minimized periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
+Bai and Breen proposed an elegant algorithm to efficiently compute the minimised periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
 However, this approach is approximate and introduces numerical error, regardless of whether particles span the periodic boundary. 
-Here, we propose an extension to the Bai and Breen algorithm removes this numerical error, allowing the correct centre of mass to be found in nearly all practical cases.
+Here, we propose an extension to the Bai and Breen algorithm that removes this numerical error, allowing the correct centre of mass to be found in nearly all practical cases.
 
@@
 This system is analogous to a fractional coordinate system. 
-A particle group of $N$ particles, can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
+A particle group of $N$ particles can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
 Each particle $i$ is then projected onto the two dimensions of a unit circle, $\bm{\xi}$ and $\bm{\zeta}$, where, 
@@
 
-Considering the limiting case of a 3 particle system, in which a central particle is moved between the 2 edge particles.
-The error in the center of mass varies, reaching its maximum very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}). 
-At the point where the particles are equally distributed, the system is perfectly symmetric and there is no numerical error in the centre of mass. 
+Considering the limiting case of a 3-particle system, where a central particle is moved between the 2 edge particles.
+The error in the centre of mass varies, reaching its maximum very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}). 
+At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no numerical error in the centre of mass. 
 Similarly, there is no error in one or two particle systems as these are inherently symmetrical. 
@@
     \includegraphics[width=8cm]{Fig2.pdf}
-    \caption{Quantification of the limiting case (bottom), a 3 particle group spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the particle group.
+    \caption{Quantification of the limiting case (bottom), a 3-particle group spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the particle group.
     This error is shown visually for a single example (top), where the Bai and Breen estimate (orange cross) is not close to the true centre of mass (green cross).}
@@
 The group of particles is repositioned to the middle of the simulation cell using the \emph{pseudo}-centre of mass, such that there is no risk that the group spans a periodic boundary condition. 
-Then, the standard weighted mean may be used to find the centre of mass, before restoring this to the original space.
+Then, the standard weighted mean may be used to find the centre of mass before restoring this to the original space.
 
-The use of this recentering assumes that the inaccurate Bai and Breen method is capable of finding a centre of mass that is close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
-This assumption holds true for all systems where the particle group is less than half the size of the periodic cell, and is often still true unless the system is very asymmetric. 
-If a group of particles spans more than half the size of the periodic box, the only way to guarantee the centre of mass is found accurately is through the inclusion of bonding or other relational information.
+The use of this recentering assumes that the inaccurate Bai and Breen method can find a centre of mass close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
+This assumption holds true for all systems where the particle group is less than half the size of the periodic cell and is often still true unless the system is very asymmetric. 
+If a group of particles spans more than half the size of the periodic box, the only way to guarantee the centre of mass is found accurately is by including bonding or other relational information.
 
@@
 To demonstrate the effectiveness of our \emph{psuedo}-centre of mass recentering extension, \cref{fig:method_comparison} compares it to the Bai and Breen method.
-To achieve this, we compute the difference between the estimated centre of mass and the true value, as a ratio of the span of the particle group. 
+To achieve this, we compute the difference between the estimated centre of mass and the true value as a ratio of the span of the particle group. 
 This error is compared with the asymmetry of the particle mass density in the periodic cell.\cite{xioajun_on_1991}
-In \cref{fig:method_comparison} as the asymmetry increases, so does the error in the Bai and Breen approach. 
+In \cref{fig:method_comparison}, as the asymmetry increases, so does the error in the Bai and Breen approach. 
 However, the \emph{psuedo}-centre of mass resampling is not affected in this way. 
-Therefore, for the majority of chemical simulations, this \emph{pseudo}-center of mass recentering extension should be used to ensure accuracy in centre of mass calculation. 
+Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering extension should be used to ensure accuracy in the centre of mass calculation. 
 %
@@
     \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
-    $2^{24}$ random configurations of one-dimensional periodic boundary spanning particle groups were generated, with between 3 and 512 particles and total length of less than half the box. 
-    For each configuration, the two approaches were used and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
-    The results have been binned along the the $x$-axis using 100 bins, spaced evenly on a log scale.}
+    $2^{24}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
+    For each configuration, the two approaches were used, and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
+    The results were binned along the $x$-axis using 100 bins spaced evenly on a log scale.}
     \label{fig:method_comparison}
@@
 
-We have presented the \emph{psuedo}-centre of mass recentering extension to the Bai and Breen approach for the determination of the centre of mass of a group of particles in a periodic cell. 
-Unlike the existing approach, this method has no error, even at very large values of asymmetry.
+We have presented the \emph{psuedo}-centre of mass recentering extension to the Bai and Breen approach for determining the centre of mass of a group of particles in a periodic cell. 
+Unlike the existing approach, this method has no error, even at very large asymmetry values.
 To enable the use of this approach, we have implemented it in the open-source Python package \textsc{kinisi}\cite{mccluskey_kinisi_2024} and hope that others will consider its use in future.
@@
 CRediT author statement: 
-H.R.: Formal analysis, Investigation, Methodology, Software, Visualization, Writing – original draft.
+H.R.: Formal analysis, Investigation, Methodology, Software, Visualisation, Writing – original draft.
 J.D.: Conceptualization, Investigation, Methodology, Supervision, Writing – review \& editing.
@@
 
-For a real valued function, the Fourier series can be described as 
+For a real-valued function, the Fourier series can be described as 
 %
@@
 %
-where, $N$ is some finite number of terms that are computed and $P$ is the period of the function. 
+where $N$ is some finite number of terms that are computed, and $P$ is the period of the function. 
 Eqn.~\ref{equ:fseries} is equivalent to
@@
 %
-where $P$ is $1$ as the Bai and Breen method projects the full box size on to a single unit circle. 
+where $P$ is $1$ as the Bai and Breen method projects the full box size onto a single unit circle. 
 To maximise the value of $f_1(x)$ a stationary point is found where,
@@
 %
-Therefore the position at which the first function of the Fourier series is maximized, $x$, can be calculated as, 
+Therefore the position at which the first function of the Fourier series is maximised, $x$, can be calculated as, 
 %

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-To demonstrate the effectiveness of our \emph{psuedo}-centre of mass recentering extension, \cref{fig:method_comparison} compares it to the Bai and Breen method.
+To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering extension, \cref{fig:method_comparison} compares it to the Bai and Breen method.
 To achieve this, we compute the difference between the estimated centre of mass and the true value as a ratio of the span of the particle group. 
@@
 In \cref{fig:method_comparison}, as the asymmetry increases, so does the error in the Bai and Breen approach. 
-However, the \emph{psuedo}-centre of mass resampling is not affected in this way. 
+However, the \emph{pseudo}-centre of mass resampling is not affected in this way. 
 Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering extension should be used to ensure accuracy in the centre of mass calculation. 
@@
     \includegraphics[width=8cm]{Fig3.pdf}
-    \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{psuedo} centre of mass recentering approach (orange line). 
+    \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{pseudo}-centre of mass recentering approach (orange line). 
     $2^{24}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
@@
 
-We have presented the \emph{psuedo}-centre of mass recentering extension to the Bai and Breen approach for determining the centre of mass of a group of particles in a periodic cell. 
+We have presented the \emph{pseudo}-centre of mass recentering extension to the Bai and Breen approach for determining the centre of mass of a group of particles in a periodic cell. 
 Unlike the existing approach, this method has no error, even at very large asymmetry values.
@@
 \begin{acknowledgements}
-We thank Richard Gowers for the insightful discussion that started us down this rabbit hole.
+We thank Richard Gowers for the insightful discussion that started us down this rabbit hole and for helpful comments on the manuscript.
 \end{acknowledgements}

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 One popular approach, described by Bai and Breen, calculates the centre of mass via the projection of the individual particles' coordinates onto a circle.
-This approach suffers from some numerical error as it is mathematically equivalent to finding the first moment of the Fourier series of the mass density. 
+This approach, which is mathematically equivalent to finding the first moment of the Fourier series of the mass density, suffers from some numerical error. 
 Here we discuss this inaccuracy and propose an extension that overcomes it, enabling improved accuracy across computational simulation. 

### 2025-01-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-Calculation of the centre of mass of some species from a periodic cell is an important aspect of chemical and physical simulation. 
+Calculation of the centre of mass of a group of particles in a periodically repeating cell is an important aspect of chemical and physical simulation. 
 One popular approach, described by Bai and Breen, calculates the centre of mass via the projection of the individual particles' coordinates onto a circle.
@@
 We thank Richard Gowers for the insightful discussion that started us down this rabbit hole and for helpful comments on the manuscript.
+We also thank Neil L. Allan for helpful discussions that lead to this work. 
+H.R. acknowledges the Engineering and Physical Sciences Research Council for funding. 
+BP plc is thanked for the financial support of J.D
+through an industrial CASE (ICASE) studentship in partnership
+with the Engineering and Physical Sciences Research Council. 
+ICASE grant reference EP/T51763X/1.
 \end{acknowledgements}

### 2025-01-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We also thank Neil L. Allan for helpful discussions that lead to this work. 
-H.R. acknowledges the Engineering and Physical Sciences Research Council for funding. 
+H.R. acknowledges the Engineering and Physical Sciences Research Council for DTP funding (EP/W524414/1). 
 BP plc is thanked for the financial support of J.D
 through an industrial CASE (ICASE) studentship in partnership
-with the Engineering and Physical Sciences Research Council. 
-ICASE grant reference EP/T51763X/1.
+with the Engineering and Physical Sciences Research Council (EP/T51763X/1).
 \end{acknowledgements}

### 2025-01-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Any group of particles will have the same number of valid centres of mass---positions in the simulation cell where the weighted relative positions of the masses sum to zero---as the number of particles (\cref{fig:problems}a).
-However, in the physical sciences, we may be concerned with the centre of mass of a single bonded molecule or a defined group of particles. 
-For a single bonded molecule, there is only one correct answer for the centre of mass.
+In the physical sciences, we may be concerned with the centre of mass of a single bonded molecule or a defined group of particles.
+For such systems, there is only one correct answer for the centre of mass.
 Practically speaking, for most simulations, it is most desirable to compute the centre of mass that minimises the periodic distance to the measured particles.
@@
 
-Bai and Breen proposed an elegant algorithm to efficiently compute the minimised periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
+Bai and Breen proposed an elegant algorithm to efficiently estimate the minimised periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
 However, this approach is approximate and introduces numerical error, regardless of whether particles span the periodic boundary. 
+For the original use case, to allow the recentering and visualisation of groups of aggregated particles, the method is effective, as numerical accuracy is less important. 
+However, within the context of computational chemistry and physics, this is not the case.  
 Here, we propose an extension to the Bai and Breen algorithm that removes this numerical error, allowing the correct centre of mass to be found in nearly all practical cases.
@@
 This system is analogous to a fractional coordinate system. 
-A particle group of $N$ particles can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
+A group of $N$ particles can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
 Each particle $i$ is then projected onto the two dimensions of a unit circle, $\bm{\xi}$ and $\bm{\zeta}$, where, 
@@
 % 
-The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses;
+The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses of $i$,
 %
@@
 Considering the limiting case of a 3-particle system, where a central particle is moved between the 2 edge particles.
-The error in the centre of mass varies, reaching its maximum very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}). 
+The error in the centre of mass varies, reaching the maximum when the central particle is very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}).
 At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no numerical error in the centre of mass. 
@@
 
-The use of this recentering assumes that the inaccurate Bai and Breen method can find a centre of mass close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
+The use of this recentering assumes that the inaccurate Bai and Breen method finds a centre of mass close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
 This assumption holds true for all systems where the particle group is less than half the size of the periodic cell and is often still true unless the system is very asymmetric. 
@@
 H.R. acknowledges the Engineering and Physical Sciences Research Council for DTP funding (EP/W524414/1). 
-BP plc is thanked for the financial support of J.D
-through an industrial CASE (ICASE) studentship in partnership
-with the Engineering and Physical Sciences Research Council (EP/T51763X/1).
+BP plc is thanked for the financial support of J.D through an industrial CASE (ICASE) studentship in partnership with the Engineering and Physical Sciences Research Council (EP/T51763X/1).
 \end{acknowledgements}

### 2025-01-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-Calculation of the centre of mass of a group of particles in a periodically repeating cell is an important aspect of chemical and physical simulation. 
+Calculation of the centre of mass of a group of particles in a periodically-repeating cell is an important aspect of chemical and physical simulation. 
 One popular approach, described by Bai and Breen, calculates the centre of mass via the projection of the individual particles' coordinates onto a circle.

### 2025-01-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \author{Andrew R. McCluskey}
-\email{andrew.mccluskey@bristol.ac.uk}
   \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
   \affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}
+\email{josh.dunn@bristol.ac.uk and andrew.mccluskey@bristol.ac.uk}
 
@@
 This approach, which is mathematically equivalent to finding the first moment of the Fourier series of the mass density, suffers from some numerical error. 
-Here we discuss this inaccuracy and propose an extension that overcomes it, enabling improved accuracy across computational simulation. 
+Here, we discuss this inaccuracy and propose an extension that overcomes it, enabling improved accuracy across computational simulation. 
 \end{abstract}

### 2025-03-19 src/scripts/Fig3.py
Updated Plotting files

@@ -1,79 +1,49 @@
 import numpy as np
+import matplotlib
 import matplotlib.pyplot as plt
-from tqdm import tqdm
-from scipy.stats import uniform
-import matplotlib as matplotlib
-from Functions import pi_centrer
+from Functions import pi_centrer,frechet_statistic,intrinsic_mean
+from Variables import font_size
+
 
 # Main
-m = int(2 ** 24) 
-k = np.zeros((m, 6))
+m = int(2 ** 19) 
+k = np.zeros((m, 4))
 x_max = 1
 x_min = 0
 
-# Variables to define wrapping
-minimum_uniform = 0.65
 
+k = np.zeros((m, 4))
 
-for i in tqdm(range(m)):
+for i in range(m):
     rng = np.random.default_rng(i)
+    minimum_uniform = rng.uniform(0,1,size=1)
     N = rng.integers(3, 512, size=1)
-    n = rng.uniform(0.1, 0.45, size=1)
+    n = rng.uniform(0.1, 0.59, size=1)
     particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
     y = particles
     particle_span = particles.max() - particles.min()
     y_pbc = y % x_max
     k[i, 0] = N
     k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
-    k[i, 2] = np.average(y)
+    k[i, 2] = intrinsic_mean(y_pbc,np.ones_like(y_pbc))
     offset = (k[i, 1] - ((x_max - x_min) / 2))
     moved = (y_pbc - offset) % (x_max - x_min)
-    k[i, 3] = np.average(moved) + offset
-    offset = (k[i, 1] - ((x_max - x_min) / 2))
-    moved = (y - offset) % (x_max - x_min)
-    # aymmetry metric
-    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
-    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
-    k[i, 4] = np.sum(np.abs(asymmetry))
-    k[i,5] = particle_span
+    k[i, 3] = (np.average(moved) + offset) % 1
 
-# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
[…]

### 2025-03-19 src/scripts/Fig4.py

@@ -1,0 +1,122 @@
+import numpy as np
+import matplotlib.gridspec as gridspec
+from Functions import test_pi_centrer
+from Variables import blue_set,cross_size,marker_set, font_size, s_set, l_set
+import matplotlib.pyplot as plt
+
+
+# Top sub figure
+coords = np.array([0.25,0.34,0.75])
+x = np.arange(0, 1, 0.0001)
+masses = np.array([1,1,1])
+
+y = np.zeros(x.shape[0])
+y[np.where(x==coords[0])] += 1
+y[np.where(x==coords[1])] += 1
+y[np.where(x==coords[2])] += 1
+
+fft = np.fft.fft(y)
+first = np.zeros_like(fft)
+first[1:2] = fft[1:2]
+
+ifft = np.fft.ifft(first).real
+
+
+# Bottom sub figure
+coords2 = np.array([0.25,0.5,0.75])
+x2 = np.arange(0, 1, 0.0001)
+masses2 = np.array([1,1,1])
+
+y2 = np.zeros(x2.shape[0])
+y2[np.where(x2==coords2[0])] += 1
+y2[np.where(x2==coords2[1])] += 1
+y2[np.where(x2==coords2[2])] += 1
+
+fft2 = np.fft.fft(y2)
+first2 = np.zeros_like(fft2)
+first2[1:2] = fft2[1:2]
+
+ifft2 = np.fft.ifft(first2).real
+
+
+# Centre of mass calcs
+True_com, Polar_com, _ = test_pi_centrer(masses,coords, box_min = 0, box_max=1)
+True_com2, Polar_com2, _ = test_pi_centrer(masses2,coords2, box_min = 0, box_max=1)
+
+height = 0.6
+particles = 3
+
+gs = gridspec.GridSpec(4, 1, width_ratios=[1], height_ratios=[1,2,1,2], wspace=0, hspace=0)
+fig = plt.figure(figsize=(4.03, 3.3))
+ax0 = fig.add_subplot(gs[0,0])
+ax1 = fig.add_subplot(gs[1,0])
+ax2 = fig.add_subplot(gs[2,0])
+ax3 = fig.add_subplot(gs[3,0])
+
+
+# Tick removal, horizontal line plotting
+y1 = 0 
+xmin = 0 
+xmax = 0.99
+
+for z in [ax0,ax2]:
+    z.set_xlim(0,1)
+    z.set_ylim(-1,1)
[…]

### 2025-03-19 src/scripts/Fig5.py

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+from scipy.stats import uniform
+import matplotlib as matplotlib
+from Functions import pi_centrer
+
+# Main
+m = int(2 ** 24) 
+k = np.zeros((m, 6))
+x_max = 1
+x_min = 0
+
+# Variables to define wrapping
+minimum_uniform = 0.65
+
+
+for i in tqdm(range(m)):
+    rng = np.random.default_rng(i)
+    N = rng.integers(3, 512, size=1)
+    n = rng.uniform(0.1, 0.45, size=1)
+    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
+    y = particles
+    particle_span = particles.max() - particles.min()
+    y_pbc = y % x_max
+    k[i, 0] = N
+    k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
+    k[i, 2] = np.average(y)
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y_pbc - offset) % (x_max - x_min)
+    k[i, 3] = np.average(moved) + offset
+    offset = (k[i, 1] - ((x_max - x_min) / 2))
+    moved = (y - offset) % (x_max - x_min)
+    # aymmetry metric
+    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
+    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
+    k[i, 4] = np.sum(np.abs(asymmetry))
+    k[i,5] = particle_span
+
+# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
+k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
+k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
+k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
+
+
+# Binning Bai and Breen error for visulation 
+N_bins = 100
+bins = np.logspace(-1, 2, N_bins)
+binned_indices = np.digitize(k[:, 4], bins)
+
+error_bb_array = np.empty([len(bins),2000000])
+error_bb_array[:] = np.nan
+err = (k[:, 1] - k[:, 2])/k[:,5]
+for z in range(0,N_bins):
[…]

### 2025-03-19 src/scripts/Functions.py

@@ -130,3 +130,22 @@ def draw_dashed_box(ax, x_min, x_max, y_min, y_max, color='black', linestyle='--
     ax.plot([x_min, x_max], [y_max, y_max], linestyle=linestyle, color=color, linewidth = linewidth)  # Top edge
     ax.plot([x_min, x_min], [y_min, y_max], linestyle=linestyle, color=color, linewidth = linewidth)  # Left edge
     ax.plot([x_max, x_max], [y_min, y_max], linestyle=linestyle, color=color, linewidth = linewidth)  # Right edge
+
+
+def frechet_statistic(theta,theta0):
+    return(np.minimum(np.abs(theta-theta0),1-np.abs(theta-theta0)))**2
+
+def intrinsic_mean(coords,masses):
+    """
+    Assumes fractional coords i.e. 0 to 1
+
+    Calculates analytical intrinsic mean via Frechet function minimisation
+
+    """
+    theta = coords #* 2 * np.pi
+    k = np.arange(len(coords))
+    ngon_means  = (np.average(theta,weights=masses) +  k/len(coords)) % 1 
+    frechet_stat = np.array([np.sum(masses*frechet_statistic(means, theta)) for means in ngon_means])
+    intrinsic_center_of_mass = ngon_means[np.where(frechet_stat==frechet_stat.min())]# / (2*np.pi)
+
+    return intrinsic_center_of_mass[0]

### 2025-03-19 src/scripts/Variables.py

@@ -6,3 +6,4 @@ font_size = 7
 cross_size = 1.5
 marker_set = 15
 l_set = 2
+s_set = 50

### 2025-03-19 src/scripts/Fig1.py
yapf and linking of plots

@@ -5,6 +5,7 @@ import matplotlib
 from matplotlib.patches import Arc
 from Functions import pi_centrer,test_pi_centrer,xi_zeta_calc,xi_zeta_bar_calc,draw_dashed_box
 from Variables import * 
+import paths
 
 
 
@@ -165,4 +166,5 @@ ax2.axis('off')
 fig.text(.30, .05, '(a)', ha='center', size = font_size)
 fig.text(.605, .05, '(b)', ha='center', size = font_size)
 fig.text(.818, .05, '(c)', ha='center', size = font_size)
-plt.show()
+plt.savefig(paths.figures / 'Fig1.pdf', bbox_inches='tight', pad_inches=0)
+plt.close()

### 2025-03-19 src/scripts/Fig2.py

@@ -11,6 +11,7 @@ import matplotlib as matplotlib
 import matplotlib.patches as patches
 from Functions import pi_centrer, xi_zeta_calc, xi_zeta_bar_calc, test_pi_centrer
 from Variables import * 
+import paths
 
 # Main
 
@@ -85,8 +86,7 @@ ax0.axvline(xrange[20]/20-0.0013, ymin = (np.abs(pi_store - MDA_store)/10)[20]/0
 ax0.axhline((np.abs(pi_store - MDA_store)/10)[20], xmax = 0.25, color = 'orange', linestyle = 'dashed', label = 'example error')
 ax0.scatter(-20,20, s=s_set, color = 'orange', marker = 'x', label = 'B&B COM',zorder=10, alpha=0.8,linewidths=cross_size)
 ax0.scatter(-20,-20, s=s_set, color = '#029E73', marker = 'x', label = 'True COM',zorder=10, alpha=0.8,linewidths=cross_size)
-
-
-
 ax0.spines['top'].set_visible(False)
-plt.show()
+
+plt.savefig(paths.figures / 'Fig2.pdf', bbox_inches='tight', pad_inches=0)
+plt.close()

### 2025-03-19 src/scripts/Fig3.py

@@ -3,6 +3,7 @@ import matplotlib
 import matplotlib.pyplot as plt
 from Functions import pi_centrer,frechet_statistic,intrinsic_mean
 from Variables import font_size
+import paths
 
 
 # Main
@@ -39,11 +40,12 @@ credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]
 axis.set_aspect('equal')
 axis.scatter(k[:,2],k[:,1],alpha = 0.4, color = '#0173B2', rasterized = True )
-axis.plot(k[:,2],k[:,3]%1, color = 'k') #linewidths = 0.001 )
+axis.plot(k[:,2],k[:,3]%1, color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')
 axis.set_xlim([0,1])
 axis.set_ylim([0,1])
 
 
-fig.savefig("method valid.pdf", bbox_inches='tight', pad_inches=0)
+plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches=0)
+plt.close()

### 2025-03-19 src/scripts/Fig4.py

@@ -3,6 +3,7 @@ import matplotlib.gridspec as gridspec
 from Functions import test_pi_centrer
 from Variables import blue_set,cross_size,marker_set, font_size, s_set, l_set
 import matplotlib.pyplot as plt
+import paths
 
 
 # Top sub figure
@@ -119,4 +120,5 @@ ax0.scatter(True_com,0, s=s_set, color = '#029E73', marker = 'x', label = 'Naive
 ax1.set_ylabel('Mass density')
 
 
-plt.show()
+plt.savefig(paths.figures / 'Fig4.pdf', bbox_inches='tight', pad_inches=0)
+plt.close()

### 2025-03-19 src/scripts/Fig5.py

@@ -4,6 +4,7 @@ from tqdm import tqdm
 from scipy.stats import uniform
 import matplotlib as matplotlib
 from Functions import pi_centrer
+import paths
 
 # Main
 m = int(2 ** 24) 
@@ -74,4 +75,6 @@ for i, ci in enumerate(credible_intervals):
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
 plt.tight_layout()
-plt.show()
+
+plt.savefig(paths.figures / 'Fig5.pdf', bbox_inches='tight', pad_inches=0)
+plt.close()

### 2025-03-20 src/tex/ms.tex

@@
 Calculation of the centre of mass of a group of particles in a periodically-repeating cell is an important aspect of chemical and physical simulation. 
-One popular approach, described by Bai and Breen, calculates the centre of mass via the projection of the individual particles' coordinates onto a circle.
+One popular approach calculates the centre of mass via the projection of the individual particles' coordinates onto a circle [Bai \& Breen, \emph{J. Graph. Tools} \textbf{13}(4), 53, (2008)].
 This approach, which is mathematically equivalent to finding the first moment of the Fourier series of the mass density, suffers from some numerical error. 

### 2025-03-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Calculation of the centre of mass of a group of particles in a periodically-repeating cell is an important aspect of chemical and physical simulation. 
-One popular approach calculates the centre of mass via the projection of the individual particles' coordinates onto a circle [Bai \& Breen, \emph{J. Graph. Tools} \textbf{13}(4), 53, (2008)].
-This approach, which is mathematically equivalent to finding the first moment of the Fourier series of the mass density, suffers from some numerical error. 
-Here, we discuss this inaccuracy and propose an extension that overcomes it, enabling improved accuracy across computational simulation. 
+One popular approach calculates the centre of mass via the projection of the individual particles' coordinates onto a circle [Bai \& Breen, \emph{J. Graph. Tools}, \textbf{13}(4), 53, (2008)].
+However, this approach involves projecting and averaging of the particles in a non-physically meaningful way resulting in inaccurate centre of masses. 
+Instead the intrinsic weighted average should be computed, but the analytical calculation of this scales as $\mathcal{O}(N\log N)$. 
+Here, we propose a more computationally efficient approach to compute the intrinsic mean suitable for the majority of chemical systems. 
 \end{abstract}
@@
   (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where two periodic cells are visible, and the dashed line indicates the particle grouping that leads to a given centre of mass (cross).
-  (b) The problem with the naive centre of mass (red cross) calculation in periodic systems, where the particle group spans a periodic boundary. 
-  (c) The projection of the particles in the bottom diagram of (b) onto a circle to find the centre of mass, which minimises the weighted distance to all particles.}
+  (b) The problem with the naïve centre of mass (red cross) calculation in periodic systems, where the particle group spans a periodic boundary. 
+  (c) The projection of the particles in the bottom diagram of (b) onto a circle to find the centre of mass (yellow cross), which minimises the weighted distance to all particles on the planar disk of the circle.}
   \label{fig:problems}
@@
 In a non-periodic system, identifying the ``correct'' centre of mass of a group of particles, the average position weighted by the masses, is trivial. 
-However, where the periodic boundary can intersect the particle group, this naive calculation may be ``incorrect'' (\cref{fig:problems}b). 
+However, where the periodic boundary can intersect the particle group, this naïve calculation may be ``incorrect'' (\cref{fig:problems}b). 
 This problem can be resolved where bonding or other relational information is present, but this is not always the case for computational simulation trajectories.
 
-Bai and Breen proposed an elegant algorithm to efficiently estimate the minimised periodic distance centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
-However, this approach is approximate and introduces numerical error, regardless of whether particles span the periodic boundary. 
+Bai and Breen proposed an algorithm to efficiently estimate the centre of mass,\cite{bai_calculating_2008} involving the projection of each orthogonal Cartesian dimension onto a circle to find the centre of mass in the periodic space (\cref{fig:problems}c).
+However, this approach appears to introduce numerical error when compared to the continuum centre of mass, regardless of whether particles span the periodic boundary. 
 For the original use case, to allow the recentering and visualisation of groups of aggregated particles, the method is effective, as numerical accuracy is less important. 
 However, within the context of computational chemistry and physics, this is not the case.  
-Here, we propose an extension to the Bai and Breen algorithm that removes this numerical error, allowing the correct centre of mass to be found in nearly all practical cases.
+Here, we discuss the nature of this error, present a general solution with non-linear scaling, and a linear-scaling solution that is accurate for the majority of chemical systems.
 
-\section{Bai and Breen Approach}
+\section{Circular Weighted Average}
 
-To outline the approach of Bai and Breen and the source of the numerical error, we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}. 
+To outline the circular averaging approach from Bai and Breen, known in directional statistics as the extrinsic mean,~\cite{hotz_intrinsic_2011} we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}.
 This system is analogous to a fractional coordinate system. 
@@
 \end{equation}
-% 
-The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses of $i$,
+The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses, $m$, of $i$,
 %
 \begin{equation}
-    \bar{\xi} = \frac{1}{N}\sum_{i=1}^{N}m_i\xi_i,
+    \bar{\xi} = \frac{\sum_{i=1}^{N}m_i\xi_i}{\sum_{i=1}^{N}m_i},
 \end{equation}
@@
 \begin{equation}
-    \bar{\zeta} = \frac{1}{N}\sum_{i=1}^{N}m_i\zeta_i.
+    \bar{\zeta} = \frac{\sum_{i=1}^{N}m_i\zeta_i}{\sum_{i=1}^{N}m_i}.
 \end{equation}
 % 
+$\bar{\xi}$ and $\bar{\zeta}$ represent the coordinates of a point on the plane formed by the circle. 
 The projection process is then reversed for the weighted average circular coordinates.
@@
 
-While not discussed in the original work, this method is equivalent to finding the first moment of the Fourier series of the mass density as a function of position (see Appendix \ref{app:math}). 
-This approach is used in signal processing to find the spectral centroid of a signal,\cite{tzanetakis_musical_2002} where it is known that it is affected by noise and asymmetry.\cite{teague_robust_2018}
-Describing an asymmetric function with a single sine wave leads to an incorrect estimate of the spectral centroid, or centre of mass.
+The approach outlined above provides the average position in a two-dimensional coordinate system, and then projects this onto a position on the circle corresponding to a position in periodic space (\cref{fig:method_comparison}c).
+However, this projection is from a non-physically meaningful position, i.e., the position does not sit on the manifold of the circle itself.
+The result of using a position on the plane of the circle, rather than the circle itself, is a discrepancy when comparing the circular weighted average centre of mass with that of the system in a non-periodic space.
 
-Considering the limiting case of a 3-particle system, where a central particle is moved between the 2 edge particles.
-The error in the centre of mass varies, reaching the maximum when the central particle is very close to the stationary particles, where the system is most asymmetrical (\cref{fig:error_quantification}).
+Consider the limiting case of a 3-particle system, where a central particle is moved between 2 edge particles, which are stationary at coordinates of 0.26 and 0.74. 
+This group of particles does not span the periodic cell and therefore the continuum centre of mass can be computed naïvely as the weighted average of the positions. 
+\cref{fig:error_quantification} presents the difference between the naïve centre of mass and that computed with the circular weighted average approach described above, normalised by the total extent of the group of particles.
+The error in the circular weighted average varies, reaching a maximum of more than \SI{30}{\percent} of the particle group size when the central particle is very close to the stationary particles, where the system is highly asymmetric.
 At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no numerical error in the centre of mass. 
-Similarly, there is no error in one or two particle systems as these are inherently symmetrical. 
+Similarly, there is no error in one or two particle systems as these are inherently symmetrical (see Appendices for further discussion of the role of asymmetry). 
 %
@@
     \includegraphics[width=8cm]{Fig2.pdf}
-    \caption{Quantification of the limiting case (bottom), a 3-particle group spanning almost exactly half the periodic box, where the solid blue line shows the error in the estimate from the Bai and Breen approach normalised by the span (length) of the particle group.
-    This error is shown visually for a single example (top), where the Bai and Breen estimate (orange cross) is not close to the true centre of mass (green cross).}
+    \caption{Quantification of the limiting case (bottom), a 3-particle group spanning close to half the periodic box, where the solid blue line shows the error between the circular weighted average centre of mass and continuum centre of mass normalised by the span (length) of the particle group.
+    This error is shown visually for a single example (top), where the circular averaged centre of mass (orange cross) is not close to the true centre of mass (green cross).}
     \label{fig:error_quantification}
@@
 
+\section{Intrinsic Weighted Average}
+
+The mean of a set of samples is the value that minimises the sum of squared distances between the mean and the samples.~\cite{neilsen_matrix_2012}
+The circular weighted average centre of mass satisfies this criterion by minimising the chordal distance between the points on the planar disk formed by the circle.~\cite{hotz_non-asymptotic_2016}
+However as outlined above, this position is not physically meaningful and requires projection onto the circular manifold, leading to the inconsistency with the continuum centre of mass.
+
+The position that matches the continuum mean of a periodic system is the intrinsic mean, which minimizes the sum of the squared arc distances between the mean and the samples. 
+As all the points exist on the periodic unit circle, all positional information is contained in their angular position. 
+This results in a computationally expensive minimisation problem~\cite{hotz_extrinsic_2013}, for which an analytical solution for the intrinsic mean can be found by producing $N$ candidate means,
+%
+\begin{equation}
+\bar{\theta}_{n} = \bar{\phi} + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
+\end{equation}
+%
+where $\bar{\phi}$ represents a naïve angular mass-weighted average position for all particles in the group.
+The true intrinsic mean is then the value in the vector, $\bar{\bm{\theta}} = [\bar{\theta}_{1}, \bar{\theta}_{2}, \ldots \bar{\theta}_{n}]$, with the lowest sum of squared arc distances between the points.
+This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_intrinsic_2011} but will always produce the true intrinsic mean.
+
+It should be noted that for bonded molecules that span over half the periodic box, the intrinsic and molecular centre of masses are not necessarily the same.
+The molecular centre of mass must be within the molecule even if this is not the minimum of the squared arc distances within the system.
+Therefore, for such systems, bonding information is required to calculate the molecular centre of mass.
+
 \section{\emph{Pseudo}-Centre of Mass Recentering}
 
-We suggest an extension that removes the error introduced by asymmetry, which involves using the Bai and Breen estimate as a \emph{pseudo}-centre of mass. 
+We propose a computationally cheap approach, with $\mathcal{O}(n)$ scaling, that finds the true continuum centre of mass. 
+This approach involves using the circular weighted averaged centre of mass as a \emph{pseudo}-centre of mass. 
 The group of particles is repositioned to the middle of the simulation cell using the \emph{pseudo}-centre of mass, such that there is no risk that the group spans a periodic boundary condition. 
-Then, the standard weighted mean may be used to find the centre of mass before restoring this to the original space.
+Then, the standard weighted average may be used to find the centre of mass before restoring this to the original space.
 
-The use of this recentering assumes that the inaccurate Bai and Breen method finds a centre of mass close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
-This assumption holds true for all systems where the particle group is less than half the size of the periodic cell and is often still true unless the system is very asymmetric. 
-If a group of particles spans more than half the size of the periodic box, the only way to guarantee the centre of mass is found accurately is by including bonding or other relational information.
+The use of this recentering assumes that the circular weighted average centre of mass method finds a value close enough to the true centre of mass that the recentering will place all particles contiguously in the periodic cell.
+This assumption holds true for all systems where the particle group is less than half the extent of the periodic cell and is often still true unless the system is very asymmetric. 
+If a group of particles spans more than half of the periodic box, the analytic intrinsic weighted average approach should be used.
 
-\section{Method Comparison}
-
-To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering extension, \cref{fig:method_comparison} compares it to the Bai and Breen method.
-To achieve this, we compute the difference between the estimated centre of mass and the true value as a ratio of the span of the particle group. 
-This error is compared with the asymmetry of the particle mass density in the periodic cell.\cite{xioajun_on_1991}
-In \cref{fig:method_comparison}, as the asymmetry increases, so does the error in the Bai and Breen approach. 
-However, the \emph{pseudo}-centre of mass resampling is not affected in this way. 
-Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering extension should be used to ensure accuracy in the centre of mass calculation. 
+To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering approach, \cref{fig:method_valid} compares it to the intrinsic weighted average centre of mass. 
+As can be seen the pseudo-centre of mass recentering approach produces a perfect estimate of the intrinsic center of mass, provided that particles do not span more half the box. 
+Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering approach should be used to ensure accuracy in the centre of mass calculation. 
 %
@@
     \centering
-    \includegraphics[width=8cm]{Fig3.pdf}
-    \caption{Error in Bai and Breen method (blue shading representing 1, 2, and 3$\sigma$ spreads) compared with the \emph{pseudo}-centre of mass recentering approach (orange line). 
-    $2^{24}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
-    For each configuration, the two approaches were used, and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
-    The results were binned along the $x$-axis using 100 bins spaced evenly on a log scale.}
-    \label{fig:method_comparison}
+    \includegraphics[width=8cm]{figures/method valid.pdf}
+    \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighte average centre of mass.
+    $2^{19}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
+    Apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the difference in the circular weighted averaging centre of mass causes the point to wrap around to the opposite side of the periodic cell.
+    }
+    \label{fig:method_valid}
     \script{Fig3.py}
@@
 
-We have presented the \emph{pseudo}-centre of mass recentering extension to the Bai and Breen approach for determining the centre of mass of a group of particles in a periodic cell. 
-Unlike the existing approach, this method has no error, even at very large asymmetry values.
+We have discussed the nature of centre of mass computation for periodic chemical simulations.
+We highlighted the problems of computing the circular weighted average centre of mass versus the intrinsic weighted average centre of mass, indicating that the latter matches the continuum centre of mass that is expected.
+However, computation of intrinsic weighted average centre of mass has non-linear scaling making it computationally inefficient for large systems. 
+To address this, we propose a linear scaling approach based on a \emph{pseudo}-centre of mass recentering that estimates the intrinsic centre of mass in the majority of chemical systems. 
 To enable the use of this approach, we have implemented it in the open-source Python package \textsc{kinisi}\cite{mccluskey_kinisi_2024} and hope that others will consider its use in future.
@@
 We also thank Neil L. Allan for helpful discussions that lead to this work. 
+We acknowledge Karol Ławniczak's insightful lectures on directional statistics that helped us to better understand this subject.
 H.R. acknowledges the Engineering and Physical Sciences Research Council for DTP funding (EP/W524414/1). 
-BP plc is thanked for the financial support of J.D through an industrial CASE (ICASE) studentship in partnership with the Engineering and Physical Sciences Research Council (EP/T51763X/1).
+BP plc is thanked for the financial support of J.D. through an industrial CASE (ICASE) studentship in partnership with the Engineering and Physical Sciences Research Council (EP/T51763X/1).
 \end{acknowledgements}
@@
 \section*{Author Contributions}
-CRediT author statement: 
 H.R.: Formal analysis, Investigation, Methodology, Software, Visualisation, Writing – original draft.
@@
 
-The data that support the findings of this study are openly available in a GitHub repository\cite{richardson_github_2025}, under an MIT license, including a complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}\cite{luger_showyourwork_2021}.
+The data that support the findings of this study are openly available in a GitHub repository,\cite{richardson_github_2025} under an MIT license, including a complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}.\cite{luger_showyourwork_2021}
 
@@
 
-\section{Mathematical Equivalency of Bai and Breen and First Moment Methods}
+\section{Mathematical Equivalency of Polar Averaging and First Moment Methods}
 \label{app:math}
 
-This appendix presents the mathematical equivalency between the Bai and Breen method for calculating the centre of mass, and finding the maximum of the first term of the Fourier series of the mass density as a function of position, known as the first moment method.\cite{teague_robust_2018}
+This appendix presents the mathematical equivalency between the circular weighted average approach to compute the centre of mass, and finding the maximum of the first term of the Fourier series of the mass density as a function of position, known as the first moment method.\cite{teague_robust_2018}
 
@@
 %
-Then $a_n$ and $b_n$ can be considered to be constants that ``weight'' the relative contributions of $\sin \left( 2\pi \dfrac{n}{P} x \right)$ and $\cos \left( 2\pi \dfrac{n}{P} x \right)$, akin to the weighted average of $\bm{\xi}$ and $\bm{\zeta}$ carried out within the B\&B method.
+Then $a_n$ and $b_n$ can be considered to be constants that ``weight'' the relative contributions of $\sin \left( 2\pi \dfrac{n}{P} x \right)$ and $\cos \left( 2\pi \dfrac{n}{P} x \right)$, akin to the weighted average of $\bm{\xi}$ and $\bm{\zeta}$ carried out within the circular weighted average approach.
 
-As stated in the main text, the Bai and Breen method is equivalent to finding the maxima of the first function of the Fourier series, $f_1$,
+As stated in the main text, the circular weighted average approach is equivalent to finding the maxima of the first function of the Fourier series, $f_1$,
 %
@@
 %
-where $P$ is $1$ as the Bai and Breen method projects the full box size onto a single unit circle. 
+where $P$ is $1$ as the circular weighted average approach projects the full box size onto a single unit circle. 
 To maximise the value of $f_1(x)$ a stationary point is found where,
@@
 %
-The negative $b_1$ and $a_1$ in the Bai and Breen method ensure that the centre of mass is always in the correct periodic box with no additional wrapping.
+The negative $b_1$ and $a_1$ in the circular weighted average approach ensure that the centre of mass is always in the correct periodic box with no additional wrapping.
+
+\section{Behaviour of Extrinsic Mean With Asymmetry}
+\label{app:fourier}
+
+As described in Appendix~\ref{app:math}, the circular weighted average centre of mass can be interpreted as identifying the maximum of the first term in the Fourier series, the first moment, of the mass density distribution.
+The first moment is the the peak position of a sine wave with a period of the periodic box length. 
+\cref{fig:sine_wave_assym} shows mass density profiles and corresponding first moment sine waves for two three-particle systems, where it can be seen that the sine wave peak tracks the position of the moving particle.
+However, this does not accurately represent the intrinsic center of mass, except in the special case where the particles are symmetrically distributed.
+
+In \cref{fig:sine_wave_assym} we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
+This is compared with a system where the particles are symmetrically positioned, in which the circular weighted averaging approach accurately estimates the continuum centre of mass. 
+By describing an asymmetric function with a single sine wave, we produce an incorrect estimate of the intrinsic centre of mass. 
+First-moment-based methods such as this are known to be sensitive to asymmetry \cite{teague_robust_2018} as the singular sine wave cannot encode asymmetrical information effectively. 
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{figures/Sine_wave.pdf}
+    \caption{Mass density histograms for two 3 particle molecules, (top) asymmetrically distributed particles [0.26, 0.34, 0.74], (bottom) symmetric molecule [0.26, 0.50, 0.74] (blue bars), and the sine waves produced from the first components of the fourier series (dashed lines).
+It can be seen in the top image that the asymmetry in the molecule causes the maximum of the sine wave to be a poor estimator of the intrinsic centre of mass.
+}
+    \label{fig:sine_wave_assym}
+\end{figure}
+%
+
+To validate and visualise the effect of asymmetry on the deviation of the circular weighted averaging centre of mass from the continuum centre of mass, the error—defined as the absolute difference between the intrinsic centre of mass and the circular averaged centre of mass, normalised by the span of the particle group was calculated.
+This error is compared with the asymmetry of the particle mass density \cite{xioajun_on_1991}. 
+In \cref{fig:method_comparison}, as the asymmetry increases, so does the difference between the circular averaged centre of mass and the intrinsic centre of mass. 
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{Fig3.pdf}
+    \caption{Error in circular averaged centre of mass method (blue shading representing 1, 2, and 3$\sigma$ spreads) as a function of asymmetry.
+    $2^{24}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
+    For each configuration, the two approaches were used, and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
+    The results were binned along the $x$-axis using 100 bins spaced evenly on a log scale.}
+    \label{fig:method_comparison}
+    \script{Fig3.py}
+\end{figure}
 

### 2025-03-20 src/scripts/Fig5.py
Move the sampling code

@@ -6,43 +6,7 @@ import matplotlib as matplotlib
 from Functions import pi_centrer
 import paths
 
-# Main
-m = int(2 ** 24) 
-k = np.zeros((m, 6))
-x_max = 1
-x_min = 0
-
-# Variables to define wrapping
-minimum_uniform = 0.65
-
-
-for i in tqdm(range(m)):
-    rng = np.random.default_rng(i)
-    N = rng.integers(3, 512, size=1)
-    n = rng.uniform(0.1, 0.45, size=1)
-    particles = rng.uniform(minimum_uniform, minimum_uniform+n, size=N)
-    y = particles
-    particle_span = particles.max() - particles.min()
-    y_pbc = y % x_max
-    k[i, 0] = N
-    k[i, 1] = pi_centrer(y_pbc, np.ones_like(y_pbc), x_min, x_max)
-    k[i, 2] = np.average(y)
-    offset = (k[i, 1] - ((x_max - x_min) / 2))
-    moved = (y_pbc - offset) % (x_max - x_min)
-    k[i, 3] = np.average(moved) + offset
-    offset = (k[i, 1] - ((x_max - x_min) / 2))
-    moved = (y - offset) % (x_max - x_min)
-    # aymmetry metric
-    y_hist, x_hist = np.histogram(moved, bins=np.linspace(x_min, x_max, 101), density=True)
-    asymmetry = y_hist[:int(y_hist.size / 2)] - y_hist[int(y_hist.size / 2):][::-1]
-    k[i, 4] = np.sum(np.abs(asymmetry))
-    k[i,5] = particle_span
-
-# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
-k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
-k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
-
+k = np.load(paths.data / 'Fig5.npy')
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-03-20 src/scripts/Fig3.py
Both 3 and 5 need the data

@@ -4,7 +4,7 @@ import matplotlib.pyplot as plt
 import matplotlib as matplotlib
 import paths
 
-k = np.loadtxt(paths.data / 'Fig3.npy')
+k = np.loadtxt(paths.data / 'Fig5.npy')
 
 # Binning Bai and Breen error for visulation
 N_bins = 100

### 2025-03-20 src/scripts/sampling.py
3 to 5

@@ -43,4 +43,4 @@ k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
 k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
 k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
 
-np.savetxt(paths.data / 'Fig3.npy', k)
+np.savetxt(paths.data / 'Fig5.npy', k)

### 2025-03-20 src/scripts/Fig4.py
Fixing figures

@@ -89,7 +89,7 @@ ax2.axvline(coords2[1], ymax = 0.5, color = 'orange')
 ax3.axvline(coords2[1], ymin = sine_wave2.max()/1.05, color = 'orange')
 
 peak_x2 = x2[y2 == 1]
-ax3.hist(peak_x2, bins=np.linspace(0, 1, 150), alpha=0.7, edgecolor='black')
+ax3.hist(peak_x2, bins=np.linspace(0, 1, 150), alpha=0.7, )
 
 ax2.scatter(Polar_com2.round(3),0, s=s_set, color = 'orange', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
 ax2.scatter(True_com2,0, s=s_set, color = '#029E73', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
@@ -113,7 +113,7 @@ ax0.axvline(coords[1], ymax = 0.5, color = 'orange')
 ax1.axvline(coords[1], ymin = sine_wave.max()/1.05, color = 'orange')
 
 peak_x = x[y == 1]
-ax1.hist(peak_x, bins=np.linspace(0, 1, 150), alpha=0.7, edgecolor='black')
+ax1.hist(peak_x, bins=np.linspace(0, 1, 150), alpha=0.7)
 
 ax0.scatter(Polar_com,0, s=s_set, color = 'orange', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)
 ax0.scatter(True_com,0, s=s_set, color = '#029E73', marker = 'x', label = 'Naive and Molecule Centre of Mass',zorder=10, alpha=0.8,linewidths=cross_size)

### 2025-03-20 src/scripts/Fig5.py

@@ -6,7 +6,7 @@ import matplotlib as matplotlib
 from Functions import pi_centrer
 import paths
 
-k = np.load(paths.data / 'Fig5.npy')
+k = np.load(paths.data / 'Fig5.npy', allow_pickle=True)
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-03-20 src/scripts/Fig5.py
Fixing figures

@@ -6,7 +6,7 @@ import matplotlib as matplotlib
 from Functions import pi_centrer
 import paths
 
-k = np.load(paths.data / 'Fig5.npy', allow_pickle=True)
+k = np.loadtxt(paths.data / 'Fig5.npy')
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-03-20 src/scripts/Fig3.py
Fixing figures

@@ -19,9 +19,12 @@ for z in range(0, N_bins):
     if length != 0:
         error_bb_array[z, 0:length] = err[np.where(binned_indices == z)[0]]
 
+# Plotting
+matplotlib.rcParams.update({'font.size': font_size})
+fig, axis = plt.subplots(figsize = (4.03, 4.03))
+
 num_bins = 100
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
-alpha = [0.6, 0.4, 0.2]
 
 axis.set_aspect('equal')
 axis.scatter(k[:,2],k[:,1],alpha = 0.4, color = '#0173B2', rasterized = True )

### 2025-03-20 src/scripts/Fig3.py
Fixing figures

@@ -2,6 +2,7 @@ import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
 import matplotlib as matplotlib
+from Variables import font_size
 import paths
 
 k = np.loadtxt(paths.data / 'Fig5.npy')

### 2025-03-20 src/scripts/Fig3.py
Update the sampling script/

@@ -5,7 +5,7 @@ import matplotlib as matplotlib
 from Variables import font_size
 import paths
 
-k = np.loadtxt(paths.data / 'Fig5.npy')
+k = np.loadtxt(paths.data / 'sampling.txt')
 
 # Binning Bai and Breen error for visulation
 N_bins = 100
@@ -28,8 +28,8 @@ num_bins = 100
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 
 axis.set_aspect('equal')
-axis.scatter(k[:,2],k[:,1],alpha = 0.4, color = '#0173B2', rasterized = True )
-axis.plot(k[:,2],k[:,3]%1, color = 'k')
+axis.scatter(k[:,2],k[:,1], alpha = 0.4, color = '#0173B2', rasterized = True )
+axis.plot(k[:,2], k[:,3] % 1, color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')
 axis.set_xlim([0,1])

### 2025-03-20 src/scripts/Fig5.py

@@ -6,7 +6,7 @@ import matplotlib as matplotlib
 from Functions import pi_centrer
 import paths
 
-k = np.loadtxt(paths.data / 'Fig5.npy')
+k = np.loadtxt(paths.data / 'sampling.txt')
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100

### 2025-03-20 src/scripts/sampling.py

@@ -1,12 +1,12 @@
 import numpy as np
 from tqdm import tqdm
 from scipy.stats import uniform
-from Functions import pi_centrer
+from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
 m = int(2**24)
-k = np.zeros((m, 6))
+k = np.zeros((m, 7))
 x_max = 1
 x_min = 0
 
@@ -26,7 +26,7 @@ for i in tqdm(range(m)):
     k[i, 2] = np.average(y)
     offset = (k[i, 1] - ((x_max - x_min) / 2))
     moved = (y_pbc - offset) % (x_max - x_min)
-    k[i, 3] = np.average(moved) + offset
+    k[i, 3] = (np.average(moved) + offset) % 1
     offset = (k[i, 1] - ((x_max - x_min) / 2))
     moved = (y - offset) % (x_max - x_min)
     # aymmetry metric
@@ -37,10 +37,11 @@ for i in tqdm(range(m)):
                                                            2):][::-1]
     k[i, 4] = np.sum(np.abs(asymmetry))
     k[i, 5] = particle_span
+    k[i, 6] = intrinsic_mean(y_pbc,np.ones_like(y_pbc))
 
 # Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
 k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
 k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
 k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
 
-np.savetxt(paths.data / 'Fig5.npy', k)
+np.savetxt(paths.data / 'sampling.txt', k)

### 2025-03-20 src/scripts/Fig3.py
Fixing figures

@@ -28,8 +28,8 @@ num_bins = 100
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 
 axis.set_aspect('equal')
-axis.scatter(k[:,2],k[:,1], alpha = 0.4, color = '#0173B2', rasterized = True )
-axis.plot(k[:,2], k[:,3] % 1, color = 'k')
+axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2', rasterized = True )
+axis.plot(k[:, 6], k[:, 3], color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')
 axis.set_xlim([0,1])

### 2025-03-20 src/scripts/sampling.py
less

@@ -5,7 +5,7 @@ from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
-m = int(2**24)
+m = int(2**18)
 k = np.zeros((m, 7))
 x_max = 1
 x_min = 0

### 2025-03-20 src/scripts/Fig3.py
Fixing figures

@@ -7,19 +7,6 @@ import paths
 
 k = np.loadtxt(paths.data / 'sampling.txt')
 
-# Binning Bai and Breen error for visulation
-N_bins = 100
-bins = np.logspace(-1, 2, N_bins)
-binned_indices = np.digitize(k[:, 4], bins)
-
-error_bb_array = np.empty([len(bins), 2000000])
-error_bb_array[:] = np.nan
-err = (k[:, 1] - k[:, 2]) / k[:, 5]
-for z in range(0, N_bins):
-    length = len(np.where(binned_indices == z)[0])
-    if length != 0:
-        error_bb_array[z, 0:length] = err[np.where(binned_indices == z)[0]]
-
 # Plotting
 matplotlib.rcParams.update({'font.size': font_size})
 fig, axis = plt.subplots(figsize = (4.03, 4.03))
@@ -32,9 +19,8 @@ axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2', rasterized = True
 axis.plot(k[:, 6], k[:, 3], color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')
-axis.set_xlim([0,1])
-axis.set_ylim([0,1])
-
+# axis.set_xlim([0,1])
+# axis.set_ylim([0,1])
 
 plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches=0)
 plt.close()

### 2025-03-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \centering
-    \includegraphics[width=8cm]{figures/method valid.pdf}
+    \includegraphics[width=8cm]{Fig3.pdf}
     \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighte average centre of mass.
-    $2^{19}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
+    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
     Apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the difference in the circular weighted averaging centre of mass causes the point to wrap around to the opposite side of the periodic cell.
@@
     \centering
-    \includegraphics[width=8cm]{figures/Sine_wave.pdf}
+    \includegraphics[width=8cm]{Fig4.pdf}
     \caption{Mass density histograms for two 3 particle molecules, (top) asymmetrically distributed particles [0.26, 0.34, 0.74], (bottom) symmetric molecule [0.26, 0.50, 0.74] (blue bars), and the sine waves produced from the first components of the fourier series (dashed lines).
@@
     \label{fig:sine_wave_assym}
+    \script{Fig4.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=8cm]{Fig3.pdf}
+    \includegraphics[width=8cm]{Fig5.pdf}
     \caption{Error in circular averaged centre of mass method (blue shading representing 1, 2, and 3$\sigma$ spreads) as a function of asymmetry.
-    $2^{24}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
+    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated (the same samples as in \cref{fig:method_valid}, with between 3 and 512 particles and a total length of less than half the box.
     For each configuration, the two approaches were used, and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.
@@
     \label{fig:method_comparison}
-    \script{Fig3.py}
+    \script{Fig5.py}
 \end{figure}

### 2025-03-20 src/scripts/Fig3.py
Fixing figures

@@ -11,16 +11,13 @@ k = np.loadtxt(paths.data / 'sampling.txt')
 matplotlib.rcParams.update({'font.size': font_size})
 fig, axis = plt.subplots(figsize = (4.03, 4.03))
 
-num_bins = 100
-credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
-
 axis.set_aspect('equal')
 axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2', rasterized = True )
 axis.plot(k[:, 6], k[:, 3], color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')
-# axis.set_xlim([0,1])
-# axis.set_ylim([0,1])
+axis.set_xlim([0,1])
+axis.set_ylim([0,1])
 
 plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches=0)
 plt.close()

### 2025-03-20 src/scripts/sampling.py

@@ -10,11 +10,9 @@ k = np.zeros((m, 7))
 x_max = 1
 x_min = 0
 
-# Variables to define wrapping
-minimum_uniform = 0.65
-
 for i in tqdm(range(m)):
     rng = np.random.default_rng(i)
+    minimum_uniform = rng.uniform(0,1,size=1)
     N = rng.integers(3, 512, size=1)
     n = rng.uniform(0.1, 0.45, size=1)
     particles = rng.uniform(minimum_uniform, minimum_uniform + n, size=N)

### 2025-03-20 src/scripts/sampling.py
Fixing figures

@@ -5,7 +5,7 @@ from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
-m = int(2**18)
+m = int(2**10)
 k = np.zeros((m, 7))
 x_max = 1
 x_min = 0

### 2025-03-20 src/scripts/sampling.py
Fixing figures

@@ -14,7 +14,7 @@ for i in tqdm(range(m)):
     rng = np.random.default_rng(i)
     minimum_uniform = rng.uniform(0,1,size=1)
     N = rng.integers(3, 512, size=1)
-    n = rng.uniform(0.1, 0.45, size=1)
+    n = rng.uniform(0.1, 0.59, size=1)
     particles = rng.uniform(minimum_uniform, minimum_uniform + n, size=N)
     y = particles
     particle_span = particles.max() - particles.min()

### 2025-03-20 src/scripts/sampling.py
Try commenting out unwrapping

@@ -38,8 +38,8 @@ for i in tqdm(range(m)):
     k[i, 6] = intrinsic_mean(y_pbc,np.ones_like(y_pbc))
 
 # Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
-k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
-k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
+# k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
+# k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
+# k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
 
 np.savetxt(paths.data / 'sampling.txt', k)

### 2025-03-20 src/scripts/Fig5.py
Try commenting out unwrapping

@@ -8,6 +8,11 @@ import paths
 
 k = np.loadtxt(paths.data / 'sampling.txt')
 
+# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
+k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
+k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
+k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
+
 # Binning Bai and Breen error for visulation 
 N_bins = 100
 bins = np.logspace(-1, 2, N_bins)

### 2025-03-20 src/scripts/sampling.py

@@ -5,7 +5,7 @@ from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
-m = int(2**10)
+m = int(2**18)
 k = np.zeros((m, 7))
 x_max = 1
 x_min = 0
@@ -36,10 +36,5 @@ for i in tqdm(range(m)):
     k[i, 4] = np.sum(np.abs(asymmetry))
     k[i, 5] = particle_span
     k[i, 6] = intrinsic_mean(y_pbc,np.ones_like(y_pbc))
-
-# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-# k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
-# k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
-# k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
 
 np.savetxt(paths.data / 'sampling.txt', k)

### 2025-03-20 src/scripts/Fig5.py
Fixing figures

@@ -8,6 +8,9 @@ import paths
 
 k = np.loadtxt(paths.data / 'sampling.txt')
 
+x_max = 1
+x_min = 0
+
 # Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
 k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
 k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max

### 2025-03-21 src/scripts/Fig3.py
Fixing broken plots

@@ -21,3 +21,5 @@ axis.set_ylim([0,1])
 
 plt.savefig(paths.figures / 'Fig3.pdf', bbox_inches='tight', pad_inches=0)
 plt.close()
+
+

### 2025-03-21 src/scripts/Fig5.py

@@ -7,14 +7,11 @@ from Functions import pi_centrer
 import paths
 
 k = np.loadtxt(paths.data / 'sampling.txt')
-
 x_max = 1
 x_min = 0
 
-# Unwrapping Centre of mass if required to allow for comparison with true Centre of mass
-k[:, 1][np.where(k[:, 1] < (x_max - x_min) * 0.5)] += x_max
-k[:, 2][np.where(k[:, 2] < (x_max - x_min) * 0.5)] += x_max
-k[:, 3][np.where(k[:, 3] < (x_max - x_min) * 0.5)] += x_max
+# Removing errors from incorrect boundary wrappings
+k = k[np.where((k[:,6] > 0.02) & (k[:,6] < 0.98))]
 
 # Binning Bai and Breen error for visulation 
 N_bins = 100
@@ -23,12 +20,11 @@ binned_indices = np.digitize(k[:, 4], bins)
 
 error_bb_array = np.empty([len(bins),2000000])
 error_bb_array[:] = np.nan
-err = (k[:, 1] - k[:, 2])/k[:,5]
+err = (k[:, 1] - k[:, 6])/k[:,5]
 for z in range(0,N_bins):
     length = len(np.where(binned_indices == z)[0])
     if length != 0:
         error_bb_array[z,0:length] = err[np.where(binned_indices == z)[0]]
-
 
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]

### 2025-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-The mean of a set of samples is the value that minimises the sum of squared distances between the mean and the samples.~\cite{neilsen_matrix_2012}
+The mean of a set of samples is the value that minimises the sum of the squared distances to the samples.~\cite{neilsen_matrix_2012}
 The circular weighted average centre of mass satisfies this criterion by minimising the chordal distance between the points on the planar disk formed by the circle.~\cite{hotz_non-asymptotic_2016}
@@
 \begin{equation}
-\bar{\theta}_{n} = \bar{\phi} + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
+\bar{x}_{n} = \bar{x}_0 + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
 \end{equation}
 %
-where $\bar{\phi}$ represents a naïve angular mass-weighted average position for all particles in the group.
-The true intrinsic mean is then the value in the vector, $\bar{\bm{\theta}} = [\bar{\theta}_{1}, \bar{\theta}_{2}, \ldots \bar{\theta}_{n}]$, with the lowest sum of squared arc distances between the points.
+where $\bar{x}_0$ represents a naïve mass-weighted average position for all particles in the group.
+The true intrinsic mean is then the value in the vector, $\bar{\bm{x}} = [\bar{x}_{1}, \bar{x}_{2}, \ldots \bar{x}_{n}]$, with the lowest sum of squared arc distances between the points. 
 This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_intrinsic_2011} but will always produce the true intrinsic mean.
@@
     \includegraphics[width=8cm]{Fig3.pdf}
-    \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighte average centre of mass.
+    \caption{A comparison of the circular weighted averaged centre of mass method (spread of blue dots) and \emph{pseudo}-centre of mass recentering approach (black line) with the intrinsic weighted average centre of mass.
     $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
@@
 
-In \cref{fig:sine_wave_assym} we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
+In \cref{fig:sine_wave_assym}, we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
 This is compared with a system where the particles are symmetrically positioned, in which the circular weighted averaging approach accurately estimates the continuum centre of mass. 

### 2025-03-21 src/scripts/Fig3.py
Testing no rasterisation

@@ -12,7 +12,7 @@ matplotlib.rcParams.update({'font.size': font_size})
 fig, axis = plt.subplots(figsize = (4.03, 4.03))
 
 axis.set_aspect('equal')
-axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2', rasterized = True )
+axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2') #, rasterized = True )
 axis.plot(k[:, 6], k[:, 3], color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')

### 2025-03-21 src/scripts/Fig3.py
test

@@ -12,7 +12,7 @@ matplotlib.rcParams.update({'font.size': font_size})
 fig, axis = plt.subplots(figsize = (4.03, 4.03))
 
 axis.set_aspect('equal')
-axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2', rasterized = True )
+axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2')#, rasterized = True )
 axis.plot(k[:, 6], k[:, 3], color = 'k')
 axis.set_xlabel('Intrinsic mean')
 axis.set_ylabel('Centre of mass')

### 2025-03-21 src/scripts/sampling.py
Try fewer samples

@@ -5,7 +5,7 @@ from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
-m = int(2**18)
+m = int(2**16)
 k = np.zeros((m, 7))
 x_max = 1
 x_min = 0

### 2025-03-21 src/scripts/sampling.py
Restore

@@ -5,7 +5,7 @@ from Functions import pi_centrer, intrinsic_mean
 import paths
 
 # Main
-m = int(2**16)
+m = int(2**18)
 k = np.zeros((m, 7))
 x_max = 1
 x_min = 0

### 2025-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The approach outlined above provides the average position in a two-dimensional coordinate system, and then projects this onto a position on the circle corresponding to a position in periodic space (\cref{fig:method_comparison}c).
-However, this projection is from a non-physically meaningful position, i.e., the position does not sit on the manifold of the circle itself.
+However, this projection is from a position with no physical meaning in the context of the periodic cell, i.e., the position does not sit on the manifold of the circle itself.
 The result of using a position on the plane of the circle, rather than the circle itself, is a discrepancy when comparing the circular weighted average centre of mass with that of the system in a non-periodic space.

### 2025-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-To outline the circular averaging approach from Bai and Breen, known in directional statistics as the extrinsic mean,~\cite{hotz_intrinsic_2011} we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}.
+To outline the circular averaging approach from Bai and Breen, known in directional statistics as the extrinsic mean,~\cite{hotz_intrinsic_2014} we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}.
 This system is analogous to a fractional coordinate system. 

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,6 +7,8 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
+matplotlib.rcParams.update({'font.size': font_size})
+
 fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
 
 # box coordinates and line

### 2025-03-24 src/scripts/Fig2.py

@@ -14,6 +14,8 @@ import paths
 
 # Main
 
+matplotlib.rcParams.update({'font.size': font_size})
+
 MDA_store = np.zeros(100)
 pi_store = np.zeros(100)
 corrected_store = np.zeros(100)

### 2025-03-24 src/scripts/Fig4.py

@@ -1,10 +1,12 @@
 import numpy as np
 import matplotlib.gridspec as gridspec
 from Functions import test_pi_centrer
-from Variables import blue_set,cross_size,marker_set, font_size, s_set, l_set
+from Variables import blue_set,cross_size, font_size, s_set, l_set
 import matplotlib.pyplot as plt
+import matplotlib
 import paths
 
+matplotlib.rcParams.update({'font.size': font_size})
 
 # Top sub figure
 coords = np.array([0.25,0.34,0.75])

### 2025-03-24 src/scripts/Fig5.py

@@ -4,7 +4,10 @@ from tqdm import tqdm
 from scipy.stats import uniform
 import matplotlib as matplotlib
 from Functions import pi_centrer
+from Variables import font_size
 import paths
+
+matplotlib.rcParams.update({'font.size': font_size})
 
 k = np.loadtxt(paths.data / 'sampling.txt')
 x_max = 1
@@ -28,7 +31,6 @@ for z in range(0,N_bins):
 
 credible_intervals = [[16, 84], [2.5, 97.5], [0.15, 99.85]]
 alpha = [0.6, 0.4, 0.2]
-matplotlib.rcParams.update({'font.size': 7})
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
 
 for i, ci in enumerate(credible_intervals):

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -9,7 +9,7 @@ import paths
 
 matplotlib.rcParams.update({'font.size': font_size})
 
-fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
+fig = plt.figure(figsize=(8.06, 3.3))
 
 # box coordinates and line
 xmin = 0

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -9,7 +9,7 @@ import paths
 
 matplotlib.rcParams.update({'font.size': font_size})
 
-fig = plt.figure(figsize=(8.06, 3.3))
+fig = plt.figure(figsize=(8.06 / 0.518, 3.3))
 
 # box coordinates and line
 xmin = 0

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,9 +7,9 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
-matplotlib.rcParams.update({'font.size': font_size})
+matplotlib.rcParams.update({'font.size': font_size * 0.518})
 
-fig = plt.figure(figsize=(8.06 / 0.518, 3.3))
+fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
 
 # box coordinates and line
 xmin = 0

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,7 +7,8 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
-matplotlib.rcParams.update({'font.size': font_size * 0.518})
+font_size *= 0.518
+matplotlib.rcParams.update({'font.size': font_size})
 
 fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
 

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,10 +7,9 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
-font_size *= 0.518
 matplotlib.rcParams.update({'font.size': font_size})
 
-fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
+fig = plt.figure(figsize=(8.06 / 0.518, 3.3))
 
 # box coordinates and line
 xmin = 0

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,9 +7,10 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
+font_size *= 0.8
 matplotlib.rcParams.update({'font.size': font_size})
 
-fig = plt.figure(figsize=(8.06 / 0.518, 3.3))
+fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))
 
 # box coordinates and line
 xmin = 0

### 2025-03-24 src/scripts/Fig1.py
font size

@@ -7,7 +7,7 @@ from Functions import pi_centrer, test_pi_centrer, xi_zeta_calc, xi_zeta_bar_cal
 from Variables import *
 import paths
 
-font_size *= 0.8
+font_size *= 0.75
 matplotlib.rcParams.update({'font.size': font_size})
 
 fig = plt.figure(figsize=(8.06, 8.06 / 4.6666667))

### 2025-03-24 src/scripts/Fig3.py
Fonts and stuff

@@ -9,7 +9,7 @@ k = np.loadtxt(paths.data / 'sampling.txt')
 
 # Plotting
 matplotlib.rcParams.update({'font.size': font_size})
-fig, axis = plt.subplots(figsize = (4.03, 4.03))
+fig, axis = plt.subplots(figsize = (3.3, 3.3))
 
 axis.set_aspect('equal')
 axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2')

### 2025-03-24 src/scripts/Fig5.py

@@ -44,6 +44,8 @@ for i, ci in enumerate(credible_intervals):
 
 axis.set_xlabel('Asymmetry')
 axis.set_ylabel("Centre of mass error / Span of particles")
+axis.set_yticks([-0.1, 0., 0.1])
+axis.set_xticks([0, 25, 50, 75, 100])
 plt.tight_layout()
 
 plt.savefig(paths.figures / 'Fig5.pdf', bbox_inches='tight', pad_inches=0)

### 2025-03-24 src/scripts/Fig3.py
restore

@@ -9,7 +9,7 @@ k = np.loadtxt(paths.data / 'sampling.txt')
 
 # Plotting
 matplotlib.rcParams.update({'font.size': font_size})
-fig, axis = plt.subplots(figsize = (3.3, 3.3))
+fig, axis = plt.subplots(figsize = (4.03, 4.03))
 
 axis.set_aspect('equal')
 axis.scatter(k[:, 6], k[:, 1], alpha = 0.4, color = '#0173B2')

### 2025-03-24 src/scripts/Fig5.py
Ylabel

@@ -43,7 +43,7 @@ for i, ci in enumerate(credible_intervals):
     
 
 axis.set_xlabel('Asymmetry')
-axis.set_ylabel("Centre of mass error / Span of particles")
+axis.set_ylabel('Normalised centre of mass error')
 axis.set_yticks([-0.1, 0., 0.1])
 axis.set_xticks([0, 25, 50, 75, 100])
 plt.tight_layout()

### 2025-03-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 For such systems, there is only one correct answer for the centre of mass.
-Practically speaking, for most simulations, it is most desirable to compute the centre of mass that minimises the periodic distance to the measured particles.
+Practically speaking, for most simulations, it is desirable to compute the centre of mass that minimises the squared periodic distance to the measured particles (Euclidean or \( L^2 \) norm ).
 %
@@
 The error in the circular weighted average varies, reaching a maximum of more than \SI{30}{\percent} of the particle group size when the central particle is very close to the stationary particles, where the system is highly asymmetric.
-At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no numerical error in the centre of mass. 
+At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no error in the centre of mass. 
 Similarly, there is no error in one or two particle systems as these are inherently symmetrical (see Appendices for further discussion of the role of asymmetry). 
@@
 The true intrinsic mean is then the value in the vector, $\bar{\bm{x}} = [\bar{x}_{1}, \bar{x}_{2}, \ldots \bar{x}_{n}]$, with the lowest sum of squared arc distances between the points. 
-This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_intrinsic_2011} but will always produce the true intrinsic mean.
+This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_extrinsic_2013} but will always produce the true intrinsic mean.
 

### 2025-04-23 src/scripts/Functions.py
Figure 6

@@ -1,12 +1,13 @@
-# Function definitions
+# Function definitions 
 import numpy as np
 import MDAnalysis as mda
 from MDAnalysis.coordinates.memory import MemoryReader
-from MDAnalysis.transformations import wrap
+from MDAnalysis.transformations import wrap 
 from MDAnalysis.transformations.boxdimensions import set_dimensions
+from geomstats.learning.frechet_mean import CircleMean
+from geomstats.geometry.hypersphere import Hypersphere
 
-
-def pi_centrer(coords, weights, box_min, box_width):
+def pi_centrer(coords,weights,box_min, box_width):
     """
     Calculates the centre of mass of a given set of coordinates and masses according to the bai and breen method.
     
@@ -18,19 +19,19 @@ def pi_centrer(coords, weights, box_min, box_width):
 
     """
     frac_coords = (coords - box_min) / box_width
-    theta = frac_coords * (2 * np.pi)
+    theta = frac_coords * (2 * np.pi) 
     xi = np.cos(theta)
     zeta = np.sin(theta)
-    xi_bar = np.average(xi, weights=weights)
-    zeta_bar = np.average(zeta, weights=weights)
+    xi_bar = np.average(xi,weights=weights)
+    zeta_bar = np.average(zeta,weights=weights)
     theta_bar = np.arctan2(-zeta_bar, -xi_bar) + np.pi
     new_s_coords = (theta_bar) / (2 * np.pi)
-    new_s_coords = new_s_coords * box_width + box_min
+    new_s_coords = new_s_coords*box_width + box_min
 
     return new_s_coords
 
 
-def xi_zeta_calc(masses, coords, box_min=0, box_width=20):
+def xi_zeta_calc(masses,coords, box_min = 0,box_width = 20):
     """
     Calculates the xi and zeta used within the bai and breen method.
     
@@ -42,14 +43,13 @@ def xi_zeta_calc(masses, coords, box_min=0, box_width=20):
 
     """
 
-    frac_coords = (coords - box_min) / box_width
+    frac_coords = (coords - box_min) / box_width 
     theta = frac_coords * (2 * np.pi) + np.pi
     xi = np.cos(theta)
     zeta = np.sin(theta)
     return xi, zeta
 
-
-def xi_zeta_bar_calc(masses, coords, box_min=0, box_width=20):
+def xi_zeta_bar_calc(masses,coords, box_min = 0, box_width = 20):
     """
[…]

### 2025-04-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 One popular approach calculates the centre of mass via the projection of the individual particles' coordinates onto a circle [Bai \& Breen, \emph{J. Graph. Tools}, \textbf{13}(4), 53, (2008)].
-However, this approach involves projecting and averaging of the particles in a non-physically meaningful way resulting in inaccurate centre of masses. 
-Instead the intrinsic weighted average should be computed, but the analytical calculation of this scales as $\mathcal{O}(N\log N)$. 
+However, this approach involves averaging of the particles in a non-physically meaningful way resulting in inaccurate centres of mass. 
+Instead the intrinsic weighted average should be computed, but the analytical calculation of this is computationally expensive and complex. 
 Here, we propose a more computationally efficient approach to compute the intrinsic mean suitable for the majority of chemical systems. 
@@
 Any group of particles will have the same number of valid centres of mass---positions in the simulation cell where the weighted relative positions of the masses sum to zero---as the number of particles (\cref{fig:problems}a).
+Computing the centre of mass will typically form part of some analysis that is performed after the simulation has been run. 
+Therefore, the simulation trajectory may involve large numbers of molecules or long trajectories, leading to a large number of individual centre of mass calculations, and efficient computation and low computational overhead is of high importance.
+
 In the physical sciences, we may be concerned with the centre of mass of a single bonded molecule or a defined group of particles.
 For such systems, there is only one correct answer for the centre of mass.
-Practically speaking, for most simulations, it is desirable to compute the centre of mass that minimises the squared periodic distance to the measured particles (Euclidean or \( L^2 \) norm ).
+Practically speaking, for most simulations, it is desirable to compute the centre of mass that minimises the squared periodic distance to the measured particles (Euclidean or \( L^2 \) norm).
 %
@@
   \caption{
-  (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where two periodic cells are visible, and the dashed line indicates the particle grouping that leads to a given centre of mass (cross).
-  (b) The problem with the naïve centre of mass (red cross) calculation in periodic systems, where the particle group spans a periodic boundary. 
+  (a) Demonstration of the $N$ possible centres of mass in an $N$ particle system, where two periodic cells are visible, and the dashed line indicates the particle grouping that leads to a given centre of mass (green cross).
+  (b) The problem with the naïve centre of mass (red cross) calculation in periodic systems; where the top shows the particle group in a single cell, compared with the bottom where the particle group spans a periodic boundary.
   (c) The projection of the particles in the bottom diagram of (b) onto a circle to find the centre of mass (yellow cross), which minimises the weighted distance to all particles on the planar disk of the circle.}
@@
 However, within the context of computational chemistry and physics, this is not the case.  
-Here, we discuss the nature of this error, present a general solution with non-linear scaling, and a linear-scaling solution that is accurate for the majority of chemical systems.
+Here, we discuss the nature of this error, present a general solution, and an optimised approach that is accurate for the majority of chemical systems.
 
@@
 
-The approach outlined above provides the average position in a two-dimensional coordinate system, and then projects this onto a position on the circle corresponding to a position in periodic space (\cref{fig:method_comparison}c).
+The approach outlined above provides the average position in a two-dimensional coordinate system, and then projects this onto a position on the circle corresponding to a position in periodic space (\cref{fig:problems}c).
 However, this projection is from a position with no physical meaning in the context of the periodic cell, i.e., the position does not sit on the manifold of the circle itself.
@@
 At the point where the particles are equally distributed, the system is perfectly symmetric, and there is no error in the centre of mass. 
-Similarly, there is no error in one or two particle systems as these are inherently symmetrical (see Appendices for further discussion of the role of asymmetry). 
+Similarly, there is no error in one or two particle systems as these are inherently symmetrical (see Appendices for further discussion of the importance of asymmetry). 
 %
@@
 \begin{equation}
-\bar{x}_{n} = \bar{x}_0 + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N.
+\bar{x}_{n} = \bar{x}_0 + \frac{n 2\pi}{N}, \quad n = 1, 2, \dots, N,
 \end{equation}
@@
 The true intrinsic mean is then the value in the vector, $\bar{\bm{x}} = [\bar{x}_{1}, \bar{x}_{2}, \ldots \bar{x}_{n}]$, with the lowest sum of squared arc distances between the points.
-This is still computationally expensive, scaling as $\mathcal{O}(N\log N)$,~\cite{hotz_extrinsic_2013} but will always produce the true intrinsic mean.
+The method above will always produce the true intrinsic mean and, therefore, the correct continuum centre of mass. 
+However, this approach has a high computational complexity and even an optimised approach is still computationally expensive with $\mathcal{O}(N\log N)$ scaling due to the need to sort the particle positions.~\cite{hotz_extrinsic_2013,miolane_geomstats_2020}.
 
@@
 
-We propose a computationally cheap approach, with $\mathcal{O}(n)$ scaling, that finds the true continuum centre of mass. 
+We propose a computationally efficient approach, that finds the true continuum centre of mass and is significantly faster than the optimised intrinsic mass computation above (see Appendix~\ref{app:complex}).
 This approach involves using the circular weighted averaged centre of mass as a \emph{pseudo}-centre of mass. 
@@
 To demonstrate the effectiveness of our \emph{pseudo}-centre of mass recentering approach, \cref{fig:method_valid} compares it to the intrinsic weighted average centre of mass. 
-As can be seen the pseudo-centre of mass recentering approach produces a perfect estimate of the intrinsic center of mass, provided that particles do not span more half the box. 
+As can be seen the pseudo-centre of mass recentering approach produces a accurate estimate of the intrinsic center of mass, provided that particles do not span more half the box. 
 Therefore, for most chemical simulations, this \emph{pseudo}-center of mass recentering approach should be used to ensure accuracy in the centre of mass calculation. 
@@
     $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated, with between 3 and 512 particles and a total length of less than half the box. 
-    Apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the difference in the circular weighted averaging centre of mass causes the point to wrap around to the opposite side of the periodic cell.
+    The apparent outliers in the top-left and bottom-right corners are not true outliers, rather, they arise when the intrinsic mean approaches the periodic boundary, the error in the circular weighted averaging centre of mass causes the estimate to wrap onto the opposite side of the periodic cell.
     }
@@
 We highlighted the problems of computing the circular weighted average centre of mass versus the intrinsic weighted average centre of mass, indicating that the latter matches the continuum centre of mass that is expected.
-However, computation of intrinsic weighted average centre of mass has non-linear scaling making it computationally inefficient for large systems. 
-To address this, we propose a linear scaling approach based on a \emph{pseudo}-centre of mass recentering that estimates the intrinsic centre of mass in the majority of chemical systems. 
+However, computation of intrinsic weighted average centre of mass is computationally complex and may be inefficient for the analysis of large trajectories. 
+To address this, we propose an efficient approach based on a \emph{pseudo}-centre of mass recentering that estimates the intrinsic centre of mass in the majority of chemical systems. 
 To enable the use of this approach, we have implemented it in the open-source Python package \textsc{kinisi}\cite{mccluskey_kinisi_2024} and hope that others will consider its use in future.
@@
 
-In \cref{fig:sine_wave_assym}, we show this sine wave for a system with particles at 0.26, 0.74, 0.34, where it is clear that a single sine wave is not capable of describing this asymmetric function. 
+In \cref{fig:sine_wave_assym}, we show this sine wave for a system with particles at 0.26,  0.34, 0.74, where it is clear that a single sine wave is not capable of describing this asymmetric function.
 This is compared with a system where the particles are symmetrically positioned, in which the circular weighted averaging approach accurately estimates the continuum centre of mass. 
 By describing an asymmetric function with a single sine wave, we produce an incorrect estimate of the intrinsic centre of mass. 
-First-moment-based methods such as this are known to be sensitive to asymmetry \cite{teague_robust_2018} as the singular sine wave cannot encode asymmetrical information effectively. 
+In this system, the extrinsic centre of mass will always be the position of the central particle, as the outer particles are almost exactly opposite each other on the circular manifold.
+Therefore, when projected onto the circular plane, the extrinsic centre of mass of the outer particles cancel each other out, leaving the total extrinsic centre of mass to be determined by the central particle.
+First-moment-based methods such as this are known to be sensitive to asymmetry \cite{teague_robust_2018} as a single sine wave cannot encode asymmetrical information effectively. 
 %
@@
 
+\section{Comparison of Computational Cost}
+\label{app:complex}
+
+The computational cost of centre of mass calculation is an important factor to allow for efficient simulation analysis. 
+Large simulations may involve many thousands or millions of particles over as many frames~\cite{lu_pflops_2021,stevens_molecular_2023} resulting in a very large number of calculations of the centre of mass of a particle group.
+As such, it is valuable to compare the computational efficiency of the two algorithms.
+
+With regards to computational scaling, the rate limiting step of the intrinsic centre of mass calculation, as implemented in \textsc{geomstats}~\cite{miolane_geomstats_2020}, is the sorting of the particle positions.
+Sorting algorithms are well documented to scale, at best, as $\mathcal{O}[N log(N)]$~\cite{knuth_sorting_1998}, where $N$ is the number of particles in the group.
+Meanwhile, all steps in our \emph{pseudo}-centre of mass recentering algorithm scale as $\mathcal{O}(N)$.
+
+However, in reality, computational scaling with the number of particles is not of great importance, as typical analyses would involve many small calculations of the centre of mass, such that constant computational overheads will dominate.
+To compare computational efficiencies, \cref{fig:comp_efficiency} presents the time taken for both algorithms with increasing number of iterations for a \num{20} particle systems.
+The \emph{pseudo}-centre of mass recentering method is approximately twice as fast as the \textsc{geomstats} implementation of the intrinsic method, due to reduced complexity and number of steps. 
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{Fig6.pdf}
+    \caption{Increasing number of calculations of the centre of mass of an example 20 particle system against computational time for both methods. 
+    All calculations run on an Apple Macbook Pro with M3 chipset.}
+    \label{fig:comp_efficiency}
+    %\script{Fig6.py}
+\end{figure}
+%
+
 \bibliography{bib}

### 2025-04-24 src/scripts/Functions.py
Remove unnecessary imports

@@ -4,8 +4,6 @@ import MDAnalysis as mda
 from MDAnalysis.coordinates.memory import MemoryReader
 from MDAnalysis.transformations import wrap 
 from MDAnalysis.transformations.boxdimensions import set_dimensions
-from geomstats.learning.frechet_mean import CircleMean
-from geomstats.geometry.hypersphere import Hypersphere
 
 def pi_centrer(coords,weights,box_min, box_width):
     """

### 2025-04-24 src/scripts/Computational_timing.py
Moving files and sorting paths

@@ -6,6 +6,7 @@ from geomstats.learning.frechet_mean import CircleMean
 
 from Functions import pi_centrer,geom_stats_intrinsic
 from Variables import font_size
+import paths
 
 times_pi = []
 ms = np.array([2**i for i in range(14, 25)])
@@ -45,5 +46,5 @@ for m in ms:
 
 time_array[:,1] = times_frech
 
-np.savetxt('comp_time.txt',time_array)
-np.savetxt('ms.txt',ms)
+np.savetxt(paths.data / 'comp_time.txt',time_array)
+np.savetxt(paths.data / 'ms.txt',ms)

### 2025-04-24 src/scripts/Fig6.py

@@ -6,8 +6,8 @@ import paths
 
 matplotlib.rcParams.update({'font.size': font_size})
 
-time_array = np.loadtxt('comp_time.txt')
-ms = np.loadtxt('ms.txt')
+time_array = np.loadtxt(paths.data / 'comp_time.txt')
+ms = np.loadtxt(paths.data / 'ms.txt')
 
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
 axis.plot(ms, time_array[:,0]/60, label = 'Pseudo-recentering',color = '#0173B2')

### 2025-04-24 src/scripts/Fig6.py
Changing axis

@@ -10,10 +10,12 @@ time_array = np.loadtxt(paths.data / 'comp_time.txt')
 ms = np.loadtxt(paths.data / 'ms.txt')
 
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
-axis.plot(ms, time_array[:,0]/60, label = 'Pseudo-recentering',color = '#0173B2')
-axis.plot(ms, time_array[:,1]/60, label = 'Intrinsic', color = 'k')
-axis.set_xlabel('Number of calculations')
-axis.set_ylabel('Time (minutes)')
+axis.plot(ms / 1e6, time_array[:,0]/60, label = 'Pseudo-recentering',color = '#0173B2')
+axis.plot(ms / 1e6, time_array[:,1]/60, label = 'Intrinsic', color = 'k')
+axis.set_xlabel('Number of calculations/$10^6$')
+axis.set_ylabel('Time/minutes')
+axis.set_xlim(0, None)
+axis.set_ylim(0, None)
 axis.legend()
 
 plt.savefig(paths.figures / 'Fig6.pdf', bbox_inches='tight', pad_inches=0)

### 2025-04-24 src/scripts/Fig6.py
Change label

@@ -10,7 +10,7 @@ time_array = np.loadtxt(paths.data / 'comp_time.txt')
 ms = np.loadtxt(paths.data / 'ms.txt')
 
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
-axis.plot(ms / 1e6, time_array[:,0]/60, label = 'Pseudo-recentering',color = '#0173B2')
+axis.plot(ms / 1e6, time_array[:,0]/60, label = 'Pseudo-centre of mass\nrecentering',color = '#0173B2')
 axis.plot(ms / 1e6, time_array[:,1]/60, label = 'Intrinsic', color = 'k')
 axis.set_xlabel('Number of calculations/$10^6$')
 axis.set_ylabel('Time/minutes')

### 2025-04-24 src/scripts/Fig6.py
Use dots

@@ -10,8 +10,8 @@ time_array = np.loadtxt(paths.data / 'comp_time.txt')
 ms = np.loadtxt(paths.data / 'ms.txt')
 
 fig, axis = plt.subplots(figsize = (4.03, 3.3))
-axis.plot(ms / 1e6, time_array[:,0]/60, label = 'Pseudo-centre of mass\nrecentering',color = '#0173B2')
-axis.plot(ms / 1e6, time_array[:,1]/60, label = 'Intrinsic', color = 'k')
+axis.plot(ms / 1e6, time_array[:,0]/60, marker='.', label = 'Pseudo-centre of\nmass recentering', color = '#0173B2')
+axis.plot(ms / 1e6, time_array[:,1]/60, marker='.', label = 'Intrinsic', color = 'k')
 axis.set_xlabel('Number of calculations/$10^6$')
 axis.set_ylabel('Time/minutes')
 axis.set_xlim(0, None)

### 2025-04-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 To compare computational efficiencies, \cref{fig:comp_efficiency} presents the time taken for both algorithms with increasing number of iterations for a \num{20} particle systems.
-The \emph{pseudo}-centre of mass recentering method is approximately twice as fast as the \textsc{geomstats} implementation of the intrinsic method, due to reduced complexity and number of steps. 
+The \emph{pseudo}-centre of mass recentering method is more than twice as fast as the \textsc{geomstats} implementation of the intrinsic method, due to reduced complexity and number of steps. 
 %
@@
     \label{fig:comp_efficiency}
-    %\script{Fig6.py}
+    \script{Fig6.py}
 \end{figure}

### 2025-04-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Computing the centre of mass will typically form part of some analysis that is performed after the simulation has been run. 
-Therefore, the simulation trajectory may involve large numbers of molecules or long trajectories, leading to a large number of individual centre of mass calculations, and efficient computation and low computational overhead is of high importance.
+Therefore, the simulation trajectory may involve large numbers of molecules or long trajectories, leading to a large number of individual centre of mass calculations.
+As such, efficient computation and low computational overhead is of high importance.
 
@@
     \caption{Error in circular averaged centre of mass method (blue shading representing 1, 2, and 3$\sigma$ spreads) as a function of asymmetry.
-    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated (the same samples as in \cref{fig:method_valid}, with between 3 and 512 particles and a total length of less than half the box.
+    $2^{18}$ random configurations of one-dimensional periodic boundary-spanning particle groups were generated (the same samples as in \cref{fig:method_valid}, with between 3 and 512 particles and a total length of less than half the box).
     For each configuration, the two approaches were used, and the asymmetry of the particle group\cite{xioajun_on_1991} is compared to the error in the estimate normalised by the span (length) of the particle group.

### 2025-04-24 src/scripts/Computational_timing.py
Add a comment

@@ -1,5 +1,9 @@
+"""
+To re run the computational timing for your own machine, install the geomstats package
+via pip and run this script from the top directory of the git repository.
+"""
+
 import numpy as np
-from tqdm import tqdm
 import time
 from geomstats.geometry.hypersphere import Hypersphere
 from geomstats.learning.frechet_mean import CircleMean
@@ -17,7 +21,7 @@ z = CircleMean(space=circle)
 
 
 times_pi = []
-for m in tqdm(ms):
+for m in ms:
     start = time.time()
     for i in range(m):
         rng = np.random.default_rng(i)
